### 2024-01-02 paper/main.tex
move overleaf to separate folder

\documentclass{article}
\usepackage{graphicx} % Required for inserting images

\title{Hierarchical Large Language Models for Multi-task problems}
\author{simen }
\date{November 2023}

\usepackage[
backend=biber,
style=alphabetic,
sorting=ynt
]{biblatex}
\usepackage{amsmath}
\addbibresource{references.bib}

\begin{document}

\maketitle

\section{Introduction}
Low-Rank  Adaption (LoRA) of an existing large-scale pre-trained Large Language Models (LLM) \cite{hu_lora_2021} has become a popular go-to approach when solving a specific text problem.

LoRA can reduce the number of trainable parameters by 10'000 times and the GPU memory requirement by 3 times \cite{hu_lora_2021}, while still outperform other fine-tuning techniques.
We are interested in LoRA adaptions for multi-task problems, such as the TULU-v2-mix \cite{ivison_camels_2023}.

For multi-task problems there are two conventional ways of training them. You can either 
\begin{enumerate}
    \item finetune each task separately
    \item train one model for all tasks (given that the textual problem contains which task is being solved for each case)
\end{enumerate}

In addition to these two approaches, we present generalization of the two approaches: the Hierarchical Low-Rank Adaption.
Given a pretrained LLM and $n$ tasks, we construct $n$ sets of LoRA parameters $\theta_n$ that each are trained on its separate task, like in approach (1) above. In additon, we create a global parameter set $\Theta$ that works as a prior to all the task specific parameters (or be regularized towards in a frequentist vocabulary).
The final posterior will look something like this:
%
\begin{equation} \label{eq:posterior}
    \sum_{n=1}^N  \sum_{i \in D_n} P(text_i | \theta_n) + \sum_{n=1}^N N(\theta_n | \Theta, \sigma)
\end{equation}

where
$D_n$ is the dataset corresponding to task $n$,
$P(text_i | \theta)$ is the likelihood of sentence $text_i$ evaluated on the parameters $\theta$,

$\sigma$ is a tuning parameter for how close the task specific parameters should be to the global parameter.

\paragraph{Why Hierarchical LLM?}
The Hierarchical LoRA Language Model will allow tasks to share information between each other through the global parameter set $\Theta$.
If a task dataset has little data, the posterior  in equation (\ref{eq:posterior}) will mainly focus on the prior, whereas if the task dataset has a lot of data, the parameters can learn more about the specific task and move away from the global "average" model.

The model above can be seen as a generalisation of the two approaches above.
As $\sigma$ goes towards zero the problem will be equal to approach (2) (assuming we have added the task in the text), 
and as $\sigma$ goes to infinity we get $n$ independent sets of parameters optimized independently as in approach (1).

\paragraph{Research Questions}
We have the following research questions
\begin{itemize}
    \item Will a maximum a posteriori hierarchical LoRA LLM  with one adapter per task outperform a single LoRA with the equivalent number of trainable parameters?
    \item Privacy: Will training a Hierarchical LoRA on multiple tasks be more privacy friendly than training on two tasks jointly?
    \item Will a posterior distribution of a LoRA LLM have "better" diversity/exploration than the sampling techniques currently used (p-sampling, nucleus samplng)
\end{itemize}

\paragraph{Evaluation}
We want to evaluate our research questions by (1) evaluating it on an open benchmark (TULU-v2-mix \cite{ivison_camels_2023}), (2) an evaluation by an GPT-4 agent, and (3) an online evaluation as a journalistic assistant in a news media article CMS, suggesting article attributes such as titles and tags.

\printbibliography

\end{document}

### 2024-01-09 src/data_utils.py
getting there on dataloading

@@ -1,0 +1,181 @@
+#%%
+from datasets import load_dataset
+import transformers
+from torch.utils.data import DataLoader
+
+def tokenize_features(sample, tokenizer, features=None, max_token_len=None):
+    if features is None:
+        features = sample.keys()
+
+    batch = {}
+    for key, val in sample.items():
+        batch[key] = f" [{key}] {val}"
+        if key in features:
+            batch[f"input_ids.{key}"] = tokenizer(
+                batch[key],
+                return_attention_mask=False,
+                add_special_tokens=False,
+                max_length=max_token_len,
+                truncation=True
+            )['input_ids']
+    return batch
+
+#dataset_nbnn = dataset_nbnn.map(lambda sample: tokenize_features(sample, tokenizer), num_proc=6)
+
+def concat_features(sample, out_config, tokenizer, max_token_len):
+    batch = {}
+    #batch = {key:val for key, val in sample.items() if "input_ids" not in key}
+
+    for config in out_config:
+        features = config['input_features'] + config['trainable_features']
+
+        tot_tokens = sum([len(sample[f"input_ids.{f}"]) for f in features])
+        excessive_tokens = tot_tokens - max_token_len +1 # +1 for eos token
+        if excessive_tokens>0:
+            sample[f"input_ids.{config['feature_to_trunc']}"] = sample[f"input_ids.{config['feature_to_trunc']}"][:-excessive_tokens]
+
+        task_name = config['task_name']
+        batch[f"input_ids.{task_name}"] = []
+        batch[f"labels.{task_name}"] = []
+        for f in config['input_features']:
+            input_col = f"input_ids.{config['task_name']}"
+            label_col = f"labels.{config['task_name']}"
+            batch[input_col] += sample[f"input_ids.{f}"]
+            batch[label_col] += [-100]*len(batch[input_col])
+
+        for tf in config['trainable_features']:
+            batch[input_col] += sample[f"input_ids.{tf}"]
+            batch[label_col] += sample[f"input_ids.{tf}"]
+        
+        # Append eos token
+        batch[input_col] += [tokenizer.eos_token_id]
+        batch[label_col] += [tokenizer.eos_token_id]
+
+    return batch
+
+def tokenize_and_concat(sample, tokenizer, out_config, max_token_len=512):
+    batch = tokenize_features(sample,tokenizer, max_token_len=max_token_len-10) # -10 to leave room for other features
[…]

### 2024-01-10 src/data_utils.py
should be all needed for data for now

@@ -1,7 +1,13 @@
 #%%
-from datasets import load_dataset
+from datasets import load_dataset, DatasetDict
+import datasets
 import transformers
 from torch.utils.data import DataLoader
+from dataclasses import dataclass
+from random import randint
+from typing import Any, Callable, Dict, List, NewType, Optional, Tuple, Union
+from transformers.utils import PaddingStrategy
+from transformers import PreTrainedTokenizerBase
 
 def tokenize_features(sample, tokenizer, features=None, max_token_len=None):
     if features is None:
@@ -11,7 +17,7 @@ def tokenize_features(sample, tokenizer, features=None, max_token_len=None):
     for key, val in sample.items():
         batch[key] = f" [{key}] {val}"
         if key in features:
-            batch[f"input_ids.{key}"] = tokenizer(
+            batch[f"{key}.input_ids"] = tokenizer(
                 batch[key],
                 return_attention_mask=False,
                 add_special_tokens=False,
@@ -20,8 +26,6 @@ def tokenize_features(sample, tokenizer, features=None, max_token_len=None):
             )['input_ids']
     return batch
 
-#dataset_nbnn = dataset_nbnn.map(lambda sample: tokenize_features(sample, tokenizer), num_proc=6)
-
 def concat_features(sample, out_config, tokenizer, max_token_len):
     batch = {}
     #batch = {key:val for key, val in sample.items() if "input_ids" not in key}
@@ -29,27 +33,31 @@ def concat_features(sample, out_config, tokenizer, max_token_len):
     for config in out_config:
         features = config['input_features'] + config['trainable_features']
 
-        tot_tokens = sum([len(sample[f"input_ids.{f}"]) for f in features])
+        tot_tokens = sum([len(sample[f"{f}.input_ids"]) for f in features])
         excessive_tokens = tot_tokens - max_token_len +1 # +1 for eos token
         if excessive_tokens>0:
-            sample[f"input_ids.{config['feature_to_trunc']}"] = sample[f"input_ids.{config['feature_to_trunc']}"][:-excessive_tokens]
+            sample[f"input_ids.{config['feature_to_trunc']}"] = sample[f"{config['feature_to_trunc']}.input_ids"][:-excessive_tokens]
 
-        task_name = config['task_name']
-        batch[f"input_ids.{task_name}"] = []
[…]

### 2024-01-12 paper/main.tex
Updates from Overleaf

@@
 
-\section{Introduction}
+\section*{Background}
 Low-Rank  Adaption (LoRA) of an existing large-scale pre-trained Large Language Models (LLM) \cite{hu_lora_2021} has become a popular go-to approach when solving a specific text problem.
@@
 
+\section{Introduction}
+\section{Related works}
+\section{Method: Hierarchical LLM}
+\section{Experiments}
+\section{Conclusion}
+
 \printbibliography

### 2024-01-14 paper/main.tex
draft a bit in the paper

@@
 \maketitle
+
+\section{Introduction}
+
+\section{Related works}
+\begin{itemize}
+    \item explain what pretraining LLM means
+    \item Explain how LORA fine tuned is done on each task
+    \item Explain how LORA finetuned also could be done on all tasks together if they have a similar data distribution.
+\end{itemize}
+
+\section{Method: Hierarchical LLM}
+
+\subsection{Task and Data Structure}
+We define our study across $D$ distinct tasks, denoted as $d=1,2,...,D$.
+For each task, our dataset comprises $N_d$ documents, with each document containing a sequence of $W_n$ tokens.
+This dataset structure is pivotal for our analysis and can be formally represented as:
+
+\begin{equation} \label{eq:data}
+\mathcal{D} = \{{ w_{d,n,1:W_n} }{n=1}^{N_d} \}_{d=1}^D
+\end{equation}
+
+In this notation, $w_{d,n,1:W_n}$ represents the token sequence in the $n^{th}$ document of the $d^{th}$ task.
+For simplification in subsequent discussions, we will omit the indices $d$ and $n$ when their reference is evident from the context.
+
+\subsection{Pretrained Autoregressive Language Model}
+Our foundational model is a pretrained autoregressive language model, denoted by its parameters $\theta_{full}$ and its architecture, referred to as $LLM$.
+The model's probability distribution is given by:
+
+\begin{equation} \label{eq:LLMprob}
+P(w_{1:W} | \theta_{full}) = \prod_{i=1}^{W-1} LLM(w_{i+1} | w_{1:i})
+\end{equation}
+%
+Here, $\theta_{full}$ represents the set of parameters obtained from pretraining on a large and diverse dataset, distinct from our current dataset $\mathcal{D}$.
+We omit to write $\theta_{full}$ explicitly in the model, as it always depdends on it.
+
+\subsection{The Hierarchical LORA finetuned}
+For each task $d$ we have a LORA parameter set $\theta_d$ that is used to fine-tune to that task.
+Hence, the likelihood of our problem follows from combining equation \ref{eq:data} and \ref{eq:LLMprob}:
+\begin{equation} \label{eq:lik}
+    \text{L}(\mathcal{D} | \theta_1, \theta_2, \ldots, \theta_D) = \prod_{d=1}^D \prod_{n=1}^{N_d} \prod_{i=1}^{W_n-1} \text{LLM}(w_{d,n,i+1} | w_{d,n,1:i}; \theta_d)
+\end{equation}
+%
+
+The prior of
+
+\begin{equation}
+    P(\theta_{1:D}) = \prod_{d=1}^D N(\theta_d ; \Theta, \sigma I)
+\end{equation}
+%
+where $\Theta$ is a set of hierarchical parameters, $I$ is a unit diagonal matrix and $\sigma \leq 0$ is a scalar hyperparameter setting how similar task parameters $\theta_d$ should be the hierarchical parameters $\Theta$.
+
+\begin{itemize}
+    \item Typically one finetune each task separately, as written in \ref{eq:lik}.
+    \item If the D tasks are more similar to eachother than the pretraining corpus, the optimal parameters $\hat{\theta}_d$ should share some information during training in order to maximize their likelihood.
+\end{itemize}
+
+\section{Experimental setup}
+We choose the `opt-350m` base model and construct three tasks that it has likely not seem as they are in Norwegian:
+\begin{itemize}
+    \item A subset of Norwegian parliament speeches of politicians containing 1000 examples
+    \item A subset of a translation dataset from Norwegian bokmål to Norwegian Nynorsk (NbAiLab-nbnn-translation) containing 1000 examples
+    \item A subset of a translation dataset from Norwegian bokmål to Norwegian Nynorsk (NbAiLab-nbnn-translation) containing 200 examples
+\end{itemize}
+
+These datasets will have a shared domain (Norwegian) that is different from what the pretrained foundational model is trained on.
+We use two tasks containing the same dataset distribution to see how sharing knowledge through the hierarchical prior can benefit datasets with few examples.
+
+\section{Conclusion}
 
@@
 \end{equation}
-
+%
 where
@@
 
-\section{Introduction}
-\section{Related works}
-\section{Method: Hierarchical LLM}
-\section{Experiments}
-\section{Conclusion}
-
 \printbibliography

### 2024-01-15 paper/main.tex
Updates from Overleaf

--- 2024-01-12 paper/main.tex
@@
 
-\section*{Background}
+\section{Background}
 Low-Rank  Adaption (LoRA) of an existing large-scale pre-trained Large Language Models (LLM) \cite{hu_lora_2021} has become a popular go-to approach when solving a specific text problem.
@@
 \section{Experiments}
-\section{Conclusion}
 

### 2024-01-27 src/data_utils.py
make token_prepend optional

@@ -37,13 +37,15 @@ def concat_features(sample, out_config, tokenizer, max_token_len):
         excessive_tokens = tot_tokens - max_token_len +1 # +1 for eos token
         if excessive_tokens>0:
             sample[f"input_ids.{config['feature_to_trunc']}"] = sample[f"{config['feature_to_trunc']}.input_ids"][:-excessive_tokens]
-
-        token_prepend = config['token_prepend']
-        batch[f"{token_prepend}.input_ids"] = []
-        batch[f"{token_prepend}.labels"] = []
+        if config['token_prepend']!="":
+            token_prepend = f"{config['token_prepend']}."
+        else:
+            token_prepend = ""
+        batch[f"{token_prepend}input_ids"] = []
+        batch[f"{token_prepend}labels"] = []
         for f in config['input_features']:
-            input_col = f"{token_prepend}.input_ids"
-            label_col = f"{token_prepend}.labels"
+            input_col = f"{token_prepend}input_ids"
+            label_col = f"{token_prepend}labels"
             batch[input_col] += sample[f"{f}.input_ids"]
             batch[label_col] += [-100]*len(sample[f"{f}.input_ids"])
 

### 2024-01-27 src/data_utils.py
add talk of norway- dataloaders

@@ -208,6 +208,96 @@ def load_noralpaca(tokenizer, max_token_len):
     dataset_noralpaca = dataset_noralpaca.map(lambda sample: tokenize_and_concat(sample, tokenizer,out_config,max_token_len=max_token_len), num_proc=None)
     return dataset_noralpaca
 
+
+def load_talkofnorway_dataset(tokenizer, max_token_len, min_token_length=10, num_tasks=3):
+    """
+    num_tasks = 3 # cut the number of speakers
+    max_token_len=512 
+    min_token_length=50
+    """
+    min_examples, max_examples = 200, 1000
+    raw = (pd.read_csv("data/ton.csv")
+        # select columns full name, party affiliation, time and text
+        .pipe(lambda df: df[['rep_name', 'text']]) # 'time','party_name', 
+        .pipe(lambda df: df.dropna())
+        # only keep speakers that have between 100 and 500 examples
+        .pipe(lambda df: df.groupby('rep_name').filter(lambda x: min_examples < len(x) < max_examples))
+        # roughly remove text that is shorter than min_token_length tokens or bigger than max_token_len tokens assuming 4 characters per token
+        .pipe(lambda df: df[df['text'].str.len() > 4*min_token_length])
+        #.pipe(lambda df: df[df['text'].str.len() < 4*max_token_len])
+        # create feature "task" to be lowercase and only letters, not even punctuation
+        .pipe(lambda df: df.assign(task=df['rep_name'].str.replace('.', '').str.replace('[^a-z]', '').str.lower()))
+        
+        # Reset index
+        .pipe(lambda df: df.reset_index(drop=True))
+        )
+    """ Raw looks like
+    time	task	party_name	text
+    0	1998-10-20T00:00:00+02:00	Sonja Irene Sjøli	Høyre	Det er en bred forståelse blant fagfolk og pol...
+    1	1998-10-20T00:00:00+02:00	Dagfinn Høybråten	Kristelig Folkeparti	Et av hovedmålene for helsetjenesten i Norge e...
+    2	1998-10-20T00:00:00+02:00	Sonja Irene Sjøli	Høyre	Jeg takker helseministeren for et fyldig og gr...
+    3	1998-10-20T00:00:00+02:00	Sonja Irene Sjøli	Høyre	
+    """
+
+    # Group by task and create train and val datasets for each
+    train_dfs = {}
+    val_dfs = {}
+    test_dfs = {}
+    task_stats = {}
+
[…]

### 2024-01-27 src/data_utils.py
import pd

@@ -8,6 +8,7 @@ from random import randint
 from typing import Any, Callable, Dict, List, NewType, Optional, Tuple, Union
 from transformers.utils import PaddingStrategy
 from transformers import PreTrainedTokenizerBase
+import pandas as pd
 
 def tokenize_features(sample, tokenizer, features=None, max_token_len=None):
     if features is None:

### 2024-02-18 src/train.py
expand to multiple optim types

@@ -26,7 +26,7 @@ class LightningHier(L.LightningModule):
         accumulate_grad_batches: int = 1,
         weight_decay:  float = 0.0,
         reg_weight: float = 0.0,
-        global_only: bool = False,
+        optim_type: bool = "joined_grad",
         *args, **kwargs
     ):
         """
@@ -43,7 +43,7 @@ class LightningHier(L.LightningModule):
         self.accumulate_grad_batches = accumulate_grad_batches
         self.weight_decay = weight_decay
         self.reg_weight = reg_weight
-        self.global_only = global_only
+        self.optim_type = optim_type
 
     def forward(self, input_ids, attention_mask, labels=None, *args, **kwargs):
         """ forward step """
@@ -93,11 +93,12 @@ class LightningHier(L.LightningModule):
                           if isinstance(val, torch.Tensor)}
 
             # Set adapter to relevant task except if global_only is True
-            if not self.global_only:
+            if self.optim_type == "global_only":
+                self.model.set_adapter("base_adapter") 
+            elif self.optim_type == "joined_grad":
+                self.model.set_adapter([task] + ["base_adapter"])
+            elif self.optim_type == "regularized":
                 self.model.set_adapter(task)
-            else:
-                self.model.set_adapter("base_adapter")
-
             output = self(
                 input_ids = batch_task['input_ids'], 
                 attention_mask = batch_task['attention_mask'], 
@@ -107,13 +108,16 @@ class LightningHier(L.LightningModule):
 
         ### hierarchical loss
         self.model.set_adapter(self.tasks + ["base_adapter"]) # Set all adapters to active again
-        
         l2_norms = self.compute_l2_norms_of_adapters()
+        
+        if self.optim_type == "regularized":
+            reg_loss = sum([val for key, val in l2_norms.items() if "l2_from_base" in key])
+        elif self.optim_type == "joined_grad":
+            reg_loss = sum([val for key, val in l2_norms.items() if ("l2/" in key) and (key != "l2/base")])
+
+        # log reg losses:
         if phase=="train":
             self.log_dict(l2_norms, on_epoch=False, sync_dist=True, on_step=True)
-
-        ## Compute total loss
[…]

### 2024-02-18 paper/main.tex
minor
add results

--- 2024-01-14 paper/main.tex
@@
 \usepackage{amsmath}
+\usepackage{booktabs}
 \addbibresource{references.bib}
@@
 \section{Related works}
+[This might be background?]
 \begin{itemize}
@@
 For each task, our dataset comprises $N_d$ documents, with each document containing a sequence of $W_n$ tokens.
-This dataset structure is pivotal for our analysis and can be formally represented as:
+This dataset structure can be formally represented as:
 
 \begin{equation} \label{eq:data}
-\mathcal{D} = \{{ w_{d,n,1:W_n} }{n=1}^{N_d} \}_{d=1}^D
+\mathcal{D} = \{ \{ w_{d,n,1:W_n} \}_{n=1}^{N_d} \}_{d=1}^D
 \end{equation}
@@
 \end{itemize}
+
+\section{Results}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{figures/results_plot.png}
+    \caption{Plot of Regularization Constant vs Perplexity}
+    \label{fig:results_plot}
+\end{figure}
+
+\begin{table}[h]
+\centering
+\caption{Empirical results of the hierarchical LORA finetuned on the three tasks. The perplexity is calculated on the validation set.}
+\label{tbl:results}
+\begin{tabular}{ccc}
+    \toprule
+    Learning Rate & Regularization & Validation Perplexity \\
+    \midrule
+        $10^{-4}$ &              0 &                 16.80 \\
+        $10^{-5}$ &       $10^{0}$ &                 16.59 \\
+        $10^{-4}$ &       $10^{1}$ &                 12.85 \\
+        $10^{-3}$ &       $10^{2}$ &                 \textbf{12.82} \\
+        $10^{-2}$ &       $10^{3}$ &                 13.26 \\
+        $10^{-1}$ &       $10^{4}$ &                 13.91 \\
+    \bottomrule
+    \end{tabular}
+\end{table}
+
+\end{document}
 

### 2024-02-18 paper/plot_generation.py

@@ -1,0 +1,53 @@
+#%%
+import matplotlib.pyplot as plt
+import seaborn as sns
+
+#%%
+import pandas as pd
+import numpy as np
+
+# Data
+data = {
+    'Learning Rate': [1e-4, 1e-5, 1e-4, 1e-3, 1e-2, 1e-1],
+    'Regularization': [0, 1, 10, 100, 1000, 10000],
+    'Validation Perplexity': [16.8, 16.59, 12.85, 12.82, 13.26, 13.91]
+}
+
+# Create DataFrame
+df = pd.DataFrame(data)
+
+# Define formatters
+formatters = {
+    'Learning Rate': lambda x: '$10^{{{}}}$'.format(int(np.log10(x))) if x != 0 else '0',
+    'Regularization': lambda x: '$10^{{{}}}$'.format(int(np.log10(x))) if x != 0 else '0',
+    'Validation Perplexity': '{:.2f}'.format
+}
+
+# Print DataFrame as LaTeX table with formatters
+print(df.to_latex(index=False, formatters=formatters, escape=False))
+#%%
+import matplotlib.pyplot as plt
+import seaborn as sns
+
+# Use seaborn style defaults and set the default figure size
+sns.set_theme(style='ticks', rc={'figure.figsize':(11.7,8.27)})
+
+# Create the plot
+plt.plot(df['Regularization'], df['Validation Perplexity'], marker='o', label='Validation Perplexity')
+
+# Add labels and title
+plt.xlabel('Regularization Constant (symlog scale)', fontsize=14)
+plt.ylabel('Perplexity', fontsize=14)
+
+# Set x-axis to symlog scale
+plt.xscale('symlog')
+
+# Add gridlines
+plt.grid(True)
+
+# Add legend
+plt.legend()
+
+# Save the plot to a file with a transparent background
+plt.savefig('figures/results_plot.png', bbox_inches='tight', dpi=300, transparent=True)
+# %%

### 2024-02-18 src/data_utils.py
add training data length per task  plotting function

@@ -316,13 +316,46 @@ def prepare_dataloaders(tokenizer, batch_size=4, max_token_len=128, *args, **kwa
         ds[phase] = datasets.concatenate_datasets([d[phase] for d in dataset_dict.values()]).shuffle(seed=42)
         dataloaders[phase] = DataLoader(ds[phase], batch_size=batch_size, collate_fn=data_collate)
     return dataloaders
+
+def build_graph_for_training_lengths():
+        import matplotlib.pyplot as plt
+        """
+        # sample of task stats
+        task_stats = {'abid q raja': {'train_len': 169, 'val_len': 84, 'test_len': 85},
+        'akhtar chaudhry': {'train_len': 110, 'val_len': 54, 'test_len': 55},
+        'aksel hagen': {'train_len': 136, 'val_len': 68, 'test_len': 67},
+        'alf egil holmelid': {'train_len': 128, 'val_len': 64, 'test_len': 63},
+        'anders b werp': {'train_len': 145, 'val_len': 72, 'test_len': 73},
+        'andré n skjelstad': {'train_len': 172, 'val_len': 86, 'test_len': 86},
+        'andré oktay dahl': {'train_len': 244, 'val_len': 122, 'test_len': 122},
+        'anette trettebergstuen': {'train_len': 120, 'val_len': 60, 'test_len': 59},
+        # plot a histogram of the number of examples per task
+    
+        """
+        # Extract speaker names and training lengths
+        speakers = list(task_stats.keys())
+        train_lengths = [stats['train_len'] for stats in task_stats.values()]
+
+        # Create bar plot
+        plt.figure(figsize=(10, 6))
+        plt.bar(speakers, train_lengths, color='steelblue', edgecolor='black')
+        plt.xlabel('Speakers', fontsize=14)
+        plt.ylabel('Training Length', fontsize=14)
+        plt.title('Training Length for each Speaker', fontsize=16)
+        plt.xticks(rotation=90, fontsize=12)  # Rotate x-axis labels for better visibility
+        plt.yticks(fontsize=12)
+        plt.grid(True)
+
+        # Save the plot as a PNG file with a transparent background
+        plt.savefig('training_lengths.png', transparent=True, bbox_inches='tight')
+
+        plt.close()
 #%%
 if __name__ == "__main__":
     tokenizer = transformers.AutoTokenizer.from_pretrained("facebook/opt-350m")
[…]

### 2024-02-18 src/train.py
restructure training slightly

@@ -306,11 +306,14 @@ def main(overwrite_params={}):
 if __name__ == "__main__":
     #main(overwrite_params={"reg_weight" : 0, "global_only" : True, "learning_rate" : 0.00001,'lora_dim' : 2,'lora_alpha' : 2})
 
-    l1 = [10, 100]
+    l1 = [0, 100]
     l2 = [1000, 1]
+    # joined grad configs
     for regloss in l1:
-        main(overwrite_params={"reg_weight" : regloss, "optim_type" : "joined_grad", "learning_rate" : 0.00001,'lora_dim' : 16,'lora_alpha' : 16})
-    #main()
+        overwrite_params={"reg_weight" : regloss, "optim_type" : "joined_grad", 'lora_dim' : 16,'lora_alpha' : 16}
+        if overwrite_params['optim_type']=="regularized":
+             overwrite_params["learning_rate"] = regloss*0.00001
+        main(overwrite_params)
 #%% RUN FOR INTERACTIVE
 """
 batch = next(iter(dataloaders['train']))

### 2024-02-18 paper/main.tex
update writing

@@
 \section{Introduction}
+Large Language Models (LLM) have become a popular tool for solving a wide range of text problems.
+LLMs are typically pretrained on a large and diverse dataset, and then fine-tuned on a specific task.
+The fine-tuning process is often done by minimizing the negative log-likelihood of the task-specific dataset, and the resulting model is then used to make predictions on new data.
+A popular fine-tuning method is Low-Rank Adaption (LoRA), which reduces the number of trainable parameters by 10'000 times and the GPU memory requirement by 3 times, while still outperforming other fine-tuning techniques (\cite{hu_lora_2021}).
+
+This paper presents a generalization of the LoRA fine-tuning method for multi-task problems, which we refer to as Hierarchical LoRA.
+The Hierarchical LoRA method allows tasks to share information between each other through a global parameter set, and can be seen as a generalization of the two conventional approaches for multi-task problems: (1) finetuning each task separately, and (2) training one model for all tasks.
+We evaluate the Hierarchical LoRA method on a benchmark dataset, and show that there exist a trade-off between the global and task-specific parameters that can be tuned to optimize the model's performance.
+
+\begin{itemize}
+    \item Typically one finetune each task separately, as written in \ref{eq:lik}.
+    \item If the D tasks are more similar to eachother than the pretraining corpus, the optimal parameters $\hat{\theta}_d$ should share some information during training in order to maximize their likelihood.
+\end{itemize}
 
 \section{Related works}
+
+\subsection{Low-Rank Adaption of Language Models}
+Low-Rank Adaption (LoRA)
+
 [This might be background?]
@@
 We define our study across $D$ distinct tasks, denoted as $d=1,2,...,D$.
-For each task, our dataset comprises $N_d$ documents, with each document containing a sequence of $W_n$ tokens.
+For each task $d$, we have a dataset $\mathcal{D}_d$ containing $N_d$ documents, with each document consisting of a sequence of $W_n$ tokens.
 This dataset structure can be formally represented as:
@@
 \begin{equation} \label{eq:data}
-\mathcal{D} = \{ \{ w_{d,n,1:W_n} \}_{n=1}^{N_d} \}_{d=1}^D
+\mathcal{D} = \{ \mathcal{D}_d \}_{d=1}^D  =  \{ \{ w_{d,n,1:W_n} \}_{n=1}^{N_d} \}_{d=1}^D
 \end{equation}
-
+%
 In this notation, $w_{d,n,1:W_n}$ represents the token sequence in the $n^{th}$ document of the $d^{th}$ task.
@@
 Here, $\theta_{full}$ represents the set of parameters obtained from pretraining on a large and diverse dataset, distinct from our current dataset $\mathcal{D}$.
-We omit to write $\theta_{full}$ explicitly in the model, as it always depdends on it.
+We omit to write $\theta_{full}$ explicitly in the model, as the autoregressive language model always depend on these parameters.
 
-\subsection{The Hierarchical LORA finetuned}
+\subsection{The Hierarchical LoRA model}
 For each task $d$ we have a LORA parameter set $\theta_d$ that is used to fine-tune to that task.
@@
 \begin{equation} \label{eq:lik}
-    \text{L}(\mathcal{D} | \theta_1, \theta_2, \ldots, \theta_D) = \prod_{d=1}^D \prod_{n=1}^{N_d} \prod_{i=1}^{W_n-1} \text{LLM}(w_{d,n,i+1} | w_{d,n,1:i}; \theta_d)
+    \text{L}(\mathcal{D} | \theta_1, \theta_2, \ldots, \theta_D) = \prod_{d=1}^D L(\mathcal{D}_d | \theta_d) = \prod_{d=1}^D \prod_{n=1}^{N_d} \prod_{i=1}^{W_n-1} \text{LLM}(w_{d,n,i+1} | w_{d,n,1:i}; \theta_d)
+\end{equation}
+%
+%
+Each task will share some information with each other, and we model this by introducing a Gaussian hierarchical prior over the task parameters $\theta_{1:D}$, given by
+\begin{equation} \label{eq:prior}
+    P(\theta_{1:D} | \Theta, \tau ) = \prod_{d=1}^D N(\theta_d ; \Theta, \frac{1}{\tau} I)
+\end{equation}
+%
+where $\Theta$ is a set of hierarchical mean parameters, $I$ is a unit diagonal matrix and $\tau \geq 0$ is a scalar precision hyperparameter controlling how similar task parameters $\theta_d$ should be to the hierarchical mean parameters $\Theta$.
+We let $\Theta$ have a uniform prior (i.e. $P(\Theta) = 1$).
+
+The posterior distribution of the hierarchical model is then given by combining equation \ref{eq:lik} and \ref{eq:prior}:
+\begin{equation} \label{eq:posterior}
+    P(\theta_{1:D} | \mathcal{D}, \tau) \propto \prod_{d=1}^D \text{L}(\mathcal{D}_d | \theta_d) \cdot P(\theta_d | \Theta, \tau)
 \end{equation}
@@
 
-The prior of
-
-\begin{equation}
-    P(\theta_{1:D}) = \prod_{d=1}^D N(\theta_d ; \Theta, \sigma I)
-\end{equation}
-%
-where $\Theta$ is a set of hierarchical parameters, $I$ is a unit diagonal matrix and $\sigma \leq 0$ is a scalar hyperparameter setting how similar task parameters $\theta_d$ should be the hierarchical parameters $\Theta$.
-
-\begin{itemize}
-    \item Typically one finetune each task separately, as written in \ref{eq:lik}.
-    \item If the D tasks are more similar to eachother than the pretraining corpus, the optimal parameters $\hat{\theta}_d$ should share some information during training in order to maximize their likelihood.
-\end{itemize}
+\subsection{Optimization}
+Given a precision hyperparameter, we want to find the maximum a posteriori (MAP) estimate of the hierarchical model.
+We will do this by using AdamW (\cite{adamW}), a gradient-based optimization algorithm, optimizing over the task parameters $\theta_{1:D}$ and the hierarchical mean parameters $\Theta$ simultaneously.
 
 \section{Experimental setup}
-We choose the `opt-350m` base model and construct three tasks that it has likely not seem as they are in Norwegian:
-\begin{itemize}
-    \item A subset of Norwegian parliament speeches of politicians containing 1000 examples
-    \item A subset of a translation dataset from Norwegian bokmål to Norwegian Nynorsk (NbAiLab-nbnn-translation) containing 1000 examples
-    \item A subset of a translation dataset from Norwegian bokmål to Norwegian Nynorsk (NbAiLab-nbnn-translation) containing 200 examples
-\end{itemize}
+% Use the Talk of Norway Dataset consisting of speeches of Norwegian parliament politicians.
+We will consider different speakers of this dataset as different tasks, and see how the hierarchical model can benefit from sharing information between these tasks.
+
+% As base model, use opt-350m as it is predominantly trained on English, and only has seen a small amount of Norwegian data from the commoncrawls dataset.
+To test our method, we optimize the model with a range of different precision hyperparameters $\tau$, and evaluate the model's perplexity on the test set for each task.
+
+We use the Talk of Norway Dataset \cite{lapponi_talk_2018}, which consists of speeches of Norwegian parliament politicians.
+Each speaker will come from different political parties and geographical areas, but the tasks will still share a common domain.
+Therefore, we can consider different speakers of this dataset as different tasks.
+Specifically, we select chronologically the 25 first speakers that have above 100 speeches in the dataset.
+This gives us a total of 25 tasks with samples ranging from 110 to 477 documents, and an average of 202 documents per task.
+The distibution of samples per task can be seen in Figure \ref{fig:task_len}.
+We split the training and test datasets randomly for each task in a 67/33 ratio.
+
+% add figure of the dataset distribution in file training_lengths.png
+\begin{figure}[h] \label{fig:task_len}
+    \centering
+    \includegraphics[width=\textwidth]{figures/training_lengths.png}
+    \caption{Number of training examples for each task}
+    \label{fig:training_lengths}
+\end{figure}
+
+We select a model that has been pretrained on a large and diverse dataset, but has not seen the specific tasks that we want to evaluate it on.
+We will use the `opt-350m` model \cite{zhang_opt_2022}, which is predominantly trained on English, and has only seen a small amount of Norwegian data from the CommonCrawls dataset.
+
+Adapting learning rate to the precision hyperparameter..
 

### 2024-02-19 paper/main.tex
add distance-to-prior graph

@@
 
+\section{Discussion}
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{figures/weights_vs_data_lengths.png}
+    \caption{The figure shows the L2-distance between each task's adapter weight and the global prior on the y-axis, and the number of training examples on the x-axis.
+We see that, as expected, tasks with more training data will have a larger distance to the global prior.}
+    \label{fig:weights_vs_datalen} 
+\end{figure}
+
 \end{document}

### 2024-02-19 src/pca_analysis.py
add pca analysis script

@@ -1,0 +1,189 @@
+#%% RUN 
+from train import *
+"""
+overwrite_params = {}
+"""
+params = {
+    'model_name': "facebook/opt-350m",
+    'max_token_length': 256,
+    'batch_size' : 64,
+    'load_in_8bit' : False,
+    'num_tasks' : 25,
+    'reg_weight' : 1.0,
+    "optim_type" : "regularized", # global_only, joined_grad , regularized
+    # LORA PARAMETERS
+    'lora_dim' : 16,
+    'lora_alpha' : 16,
+    'lora_dropout' : 0.0,
+    'lora_target_modules' : None,
+    # OPTIM PARAMETERS
+    'learning_rate' : 0.0001,
+    'weight_decay' : 0,
+    'accumulate_grad_batches' : 1,
+    'early_stopping_patience_epochs' : 8,
+    'max_epochs' : 1000,
+    'precision': 32,
+    'log_every_n_steps' : 10,
+    'val_check_interval' : 1.0, # 0.25 = 4 times per epoch
+}
+
+tokenizer = load_tokenizer(params)
+dataloaders, task_stats = data_utils.prepare_talkofnorway_dataloaders(tokenizer, **params)
+
+pl_model = load_model(params, task_stats=task_stats, tokenizer=tokenizer, checkpoint_path="/home/simen.eide@schibsted.com/hier-llm/src/logs_regularized/1kepoch-reg:10-lr:0.0001-global:False/version_1/checkpoints/epoch=281-step=22560.ckpt")
+
+model = pl_model.model
+#%% concat all adapter weights from different tasks
+parameter_store = {key : [] for key in task_stats.keys()}
+parameter_away_from_base_store = {key : [] for key in task_stats.keys()}
+for base_key, val in model.named_parameters():
+    if "base_adapter" in base_key:
+        for adapter in task_stats.keys():
+            adapter_key = base_key.replace("base_adapter", adapter)
+            par = model.get_parameter(adapter_key).data
+            # Add to list
+            parameter_store[adapter].append( par.flatten() )
+            # Compare to base and add to list
+            base_par = model.get_parameter(base_key).data
+            parameter_away_from_base_store[adapter].append( (par - base_par).flatten() )
+
+adapter_weights = torch.stack([torch.concat(parameter_store[task]) for task in task_stats.keys()])
+
[…]

### 2024-03-04 paper/main.tex
more writing on lora++

@@
 \usepackage{amsmath}
+\usepackage{amssymb}
+\newcommand{\R}{\mathbb{R}}
 \usepackage{booktabs}
@@
 \section{Related works}
+
+\subsection{Multi-task Learning}
+Multi-task Learning \cite{caruana_multitask_nodate} is a popular machine learning approach for training one model on multiple tasks.
+A common approach is to share the lower layers of a neural network between tasks, and have separate output layers for each task.
+Our paper focuses on a different approach, where we share information between tasks through a hierarchical prior over the task-specific parameters.
 
@@
 \subsection{Pretrained Autoregressive Language Model}
-Our foundational model is a pretrained autoregressive language model, denoted by its parameters $\theta_{full}$ and its architecture, referred to as $LLM$.
+Our foundational model is a pretrained autoregressive language model, denoted by its parameters $\theta_{full}$ and its architecture, referred to as a Large Language Model ($LLM$).
+The parameters $\theta_{full}$ has been obtained from pretraining on a large and diverse dataset, distinct from our current dataset $\mathcal{D}$.
+The model's probability distribution is given by:
 The model's probability distribution is given by:
@@
 %
-Here, $\theta_{full}$ represents the set of parameters obtained from pretraining on a large and diverse dataset, distinct from our current dataset $\mathcal{D}$.
+where $LLM(w_{i+1} | w_{1:i})$ is the probability of the next token $w_{i+1}$ given the previous tokens $w_{1:i}$ in the document.
 We omit to write $\theta_{full}$ explicitly in the model, as the autoregressive language model always depend on these parameters.
 
+\subsection{Low-Rank Adaption (LoRA)}
+A Large Language Model contains many linear layers, and the LoRA method reduces the number of trainable parameters by reparameterizing the weight parameters ($W \in \R^{n_1 \times n_2}$) of these layers into the original pretrained weights $W_{full}$ and a low-rank decomposition $A$ and $B$ (\cite{hayou_lora_2024}):
+\begin{equation} \label{eq:LoRA}
+    W = W_{full} + \frac{\alpha}{r} BA 
+\end{equation}
+%
+where $W_{full} \in \R^{n_1 \times n_2}$ is the original pretrained weight, $B \in \R^{n_1 \times r}$ and $A \in \R^{r \times n_2}$ are the low-rank factors, $r \ll \text{min}(n_1, n_2)$ is the low rank dimension and $\alpha$ is a scaling hyperparameter.
+
+During LoRA fine-tuning, the low-rank factors $A$ and $B$ are learned as normal parameters, while the original pretrained weights $W_{full}$ are kept fixed.
+
 \subsection{The Hierarchical LoRA model}
-For each task $d$ we have a LORA parameter set $\theta_d$ that is used to fine-tune to that task.
-Hence, the likelihood of our problem follows from combining equation \ref{eq:data} and \ref{eq:LLMprob}:
+In the hierarchical LoRA model, we introduce a LoRA parameter set for each task $d$, denoted as $\theta_d$, containing all the low-rank decomposition matricies $A$ and $B$ from each decomposed layer in the Large Language Model.
+The likelihood of the data $\mathcal{D}$ given the task parameters $\theta_{1:D}$ is then given by (combining equation \ref{eq:data} and \ref{eq:LLMprob}):
 \begin{equation} \label{eq:lik}

### 2024-03-11 paper/main.tex
Updates from Overleaf

@@
 
-\subsection{Low-Rank Adaption of Language Models}
-Low-Rank Adaption (LoRA)
+\subsection{Finetuning methods}
+Talk about alternative finetuning methods to lora: finetune top layer (traditional),
 
-[This might be background?]
-\begin{itemize}
-    \item explain what pretraining LLM means
-    \item Explain how LORA fine tuned is done on each task
-    \item Explain how LORA finetuned also could be done on all tasks together if they have a similar data distribution.
-\end{itemize}
+sira (\cite{zhu_sira_2023})
 
@@
 
-During LoRA fine-tuning, the low-rank factors $A$ and $B$ are learned as normal parameters, while the original pretrained weights $W_{full}$ are kept fixed.
+During LoRA fine-tuning, the low-rank factors $A$ and $B$ are learned from data as normal parameter matricies, while the original pretrained weights $W_{full}$ are kept fixed.
 
@@
 
-\end{document}
-
-These datasets will have a shared domain (Norwegian) that is different from what the pretrained foundational model is trained on.
-We use two tasks containing the same dataset distribution to see how sharing knowledge through the hierarchical prior can benefit datasets with few examples.
+%\end{document}
 

### 2024-03-29 paper/main.tex
writing

@@
 
-\title{Hierarchical Large Language Models for Multi-task problems}
+\title{HLoRA: Hierarchical Large Language Models for Multi-task problems}
 \author{simen }
@@
 \usepackage{booktabs}
-\addbibresource{references.bib}
+\addbibresource{My Library.bib}
 
@@
 \subsection{Finetuning methods}
-Talk about alternative finetuning methods to lora: finetune top layer (traditional),
+A popular approach to finetuning neural networks involves adjusting only the top layer (or the n top layers) of a pre-trained neural network (e.g.
+\cite{yosinski2014transferable}).
+These techniques observe that features learned in the lower layers of the network are general and transferable, while the top layers are more task-specific.
+This approach has been widely adopted in neural networks.
+However, we focus on the LoRA method, which allows us to train a low rank decomposition of the entire network.
 
-sira (\cite{zhu_sira_2023})
+There are several variations of the LoRA algorithm.
+The Sparse Mixture of Low Rank Adapters (SIRA) \cite{zhu_sira_2023}, for instance, creates a Mixture of Expert on the low-rank parameters.
+Another variation, the Weight Decomposed Low-Rank Adaption (DoRA) \cite{hayou_lora_2024}, decomposes the low-rank parameters into scale and unit norm direction parameters.
+There's also the Efficient Low Rank Adaption (LoRA+) \cite{hayou_lora_2024}, which improves the existing LoRA optimization algorithm by adjusting the learning rate of the low-rank parameters.
+All these methods are designed to improve the performance of the LoRA algorithm, and can be used in conjunction with the Hierarchical LoRA method.
 
 \section{Method: Hierarchical LLM}
+This section describes the Hierarchical LoRA method for multi-task problems.
+We start by defining the task and data structure, and then introduce the pretrained autoregressive language model.
+We then describe the LoRA method, and finally present our Hierarchical LoRA model.
 
@@
 Given a precision hyperparameter, we want to find the maximum a posteriori (MAP) estimate of the hierarchical model. 
-We will do this by using AdamW (\cite{adamW}), a gradient-based optimization algorithm, optimizing over the task parameters $\theta_{1:D}$ and the hierarchical mean parameters $\Theta$ simultaneously.
+We will do this by using AdamW (\cite{adamW}), a gradient-based optimization algorithm, optimizing over the task parameters $\theta_{1:D}$ and the hierarchical mean parameters $\Theta$ simultaneously optimizing the likelihood and the prior in equation \ref{eq:posterior}.
+This implies that the task parameters with have gradients both towards minimizing the convential LLM loss and the hierarchical mean parameters.
+The hierarchical mean parameters $\Theta$ will have gradients towards the average of the task parameters, and learn to represent the global structure of the tasks after the task parameters have been partly optimized.
+
+An important observation is that the gradient of the prior term in equation \ref{eq:posterior} is proportional to the precision hyperparameter $\tau$.
+This means that the optimization will more steps to converge for larger values of $\tau$, as the task parameters will be more constrained towards the hierarchical mean parameters.
+Therefore, we adjust the learning rate of the AdamW optimizer to be proportional to the precision hyperparameter $\tau$, and we find in informal tests that this cause the optimizations to converge at a similar rate for different values of $\tau$.
 
@@
 The distibution of samples per task can be seen in Figure \ref{fig:task_len}.
-We split the training and test datasets randomly for each task in a 67/33 ratio.
+We leave 33\% of the data for the test set.
 
@@
 \section{Results}
+The results of the hierarchical LoRA model on the Talk of Norway Dataset are shown in Table \ref{tbl:results} and Figure \ref{fig:results_plot}.
+We see that the model achieves the best validation perplexity of 12.82 when using a learning rate of $10^{-3}$ and a regularization constant of $10^{2}$.
+This indicates that the model benefits from sharing information between the tasks, and that the global prior helps to improve the performance of the individual task parameters.
 
@@
 \centering
-\caption{Empirical results of the hierarchical LORA finetuned on the three tasks. The perplexity is calculated on the validation set.}
+\caption{Empirical results of the hierarchical LORA finetuned on the three tasks. The perplexity is calculated on the test set.}
 \label{tbl:results}
@@
 \section{Discussion}
+
+% Discuss overall result
+% Fetch perplexity for each task and discuss how it improves results significantly for some tasks (less data?), and less larger data tasks.
+% Give some advice on how to use this method in practice.
+
 \begin{figure}[h]

### 2024-03-29 src/generate_plots.py
add figure plot and figures

@@ -1,0 +1,164 @@
+#%% data
+import numpy as np
+import pandas as pd
+from tbparse import SummaryReader
+import numpy as np
+import seaborn as sns
+import matplotlib.pyplot as plt
+
+def build_dataset_of_dataset_sizes():
+    # Gathered from the dataloader but pasted here for convenience
+    task_stats = {'abid q raja': {'train_len': 169, 'val_len': 84, 'test_len': 85},
+    'akhtar chaudhry': {'train_len': 110, 'val_len': 54, 'test_len': 55},
+    'aksel hagen': {'train_len': 136, 'val_len': 68, 'test_len': 67},
+    'alf egil holmelid': {'train_len': 128, 'val_len': 64, 'test_len': 63},
+    'anders b werp': {'train_len': 145, 'val_len': 72, 'test_len': 73},
+    'andré n skjelstad': {'train_len': 172, 'val_len': 86, 'test_len': 86},
+    'andré oktay dahl': {'train_len': 244, 'val_len': 122, 'test_len': 122},
+    'anette trettebergstuen': {'train_len': 120, 'val_len': 60, 'test_len': 59},
+    'anita apelthun sæle': {'train_len': 128, 'val_len': 64, 'test_len': 63},
+    'anne tingelstad wøien': {'train_len': 181, 'val_len': 90, 'test_len': 91},
+    'anniken huitfeldt': {'train_len': 384, 'val_len': 192, 'test_len': 192},
+    'ansgar gabrielsen': {'train_len': 404, 'val_len': 202, 'test_len': 203},
+    'arild grande': {'train_len': 160, 'val_len': 80, 'test_len': 81},
+    'arne lyngstad': {'train_len': 125, 'val_len': 62, 'test_len': 63},
+    'arve kambe': {'train_len': 158, 'val_len': 78, 'test_len': 79},
+    'asmund kristoffersen': {'train_len': 152, 'val_len': 76, 'test_len': 77},
+    'audun lysbakken': {'train_len': 452, 'val_len': 226, 'test_len': 227},
+    'bendiks h arnesen': {'train_len': 238, 'val_len': 118, 'test_len': 119},
+    'bente thorsen': {'train_len': 147, 'val_len': 74, 'test_len': 73},
+    'bjørg tørresdal': {'train_len': 214, 'val_len': 106, 'test_len': 107},
[…]

### 2024-03-31 paper/main.tex
lots of writing and plots

@@
 \usepackage{graphicx} % Required for inserting images
+\usepackage{authblk}
 
-\title{HLoRA: Hierarchical Large Language Models for Multi-task problems}
-\author{simen }
-\date{November 2023}
+\title{HLoRA: Hierarchical Low-Rank Adaption for Multi-task Large Language Models}
+\author[1]{Simen Eide}
+\author[2]{Arnoldo Frigessi}
+\affil[1]{University of Oslo, Schibsted}
+\affil[1]{University of Oslo}
+\date{March 2024}
 
@@
 \maketitle
+\abstract{Lorem Ipsum}
 
 \section{Introduction}
-Large Language Models (LLM) have become a popular tool for solving a wide range of text problems.
+Large Language Models (LLM) have become a popular model architechture for solving a wide range of text problems.
 LLMs are typically pretrained on a large and diverse dataset, and then fine-tuned on a specific task.
-The fine-tuning process is often done by minimizing the negative log-likelihood of the task-specific dataset, and the resulting model is then used to make predictions on new data.
-A popular fine-tuning method is Low-Rank Adaption (LoRA), which reduces the number of trainable parameters by 10'000 times and the GPU memory requirement by 3 times, while still outperforming other fine-tuning techniques (\cite{hu_lora_2021}).
+The fine-tuning process is often done by maximizing the next token log likelihood of the task-specific dataset, and the resulting model is then used to make predictions and generations on new data.
+A popular fine-tuning method is Low-Rank Adaption (LoRA), which reduces the number of trainable parameters by up to 10'000 times and the GPU memory requirement by 3 times, while still outperforming other fine-tuning techniques (\cite{hu_lora_2021}).
 
-This paper presents a generalization of the LoRA fine-tuning method for multi-task problems, which we refer to as Hierarchical LoRA.
-The Hierarchical LoRA method allows tasks to share information between each other through a global parameter set, and can be seen as a generalization of the two conventional approaches for multi-task problems: (1) finetuning each task separately, and (2) training one model for all tasks.
-We evaluate the Hierarchical LoRA method on a benchmark dataset, and show that there exist a trade-off between the global and task-specific parameters that can be tuned to optimize the model's performance.
+When applying fine-tuning of LLMs to problems, the practitioner often wants to solve multiple similar tasks at the same time. 
+For instance, a news organization might want to use an LLM to generate various textual elements for a given article such as headlines, keywords, and summaries, all based on the main body of the text.
+This would involve fine-tuning the model for each of these distinct but related tasks.
+Similarly, a customer service chatbot might need to be fine-tuned to handle inquiries across a range of topics, such as product information, order status, and troubleshooting.
+Additionally, the chatbot might need to adapt its responses based on the user's tone or the formality of the conversation, requiring further fine-tuning.
+In a more creative context, a scriptwriting AI might be fine-tuned to generate dialogues for different characters in a movie, each with their unique style and vocabulary.
+This would involve fine-tuning the model for each character, a set of related but distinct tasks.
 
-\begin{itemize}
-    \item Typically one finetune each task separately, as written in \ref{eq:lik}.
-    \item If the D tasks are more similar to eachother than the pretraining corpus, the optimal parameters $\hat{\theta}_d$ should share some information during training in order to maximize their likelihood.
-\end{itemize}
+In these cases, the practitioner can either (1) train one model for each task separately, or (2) train one model for all tasks and condition the specific output wanted (e.g \cite{Raffel2019}).
+There is a trade-off between these two approaches: Train each task separately and the model will be able to specialize on each task, but it may suffer from limited data for each task.
+Train one model for all tasks, and the model will be able to share information between tasks, but it may not be able to specialize on each task due to limited model capacity.
 
-\section{Related works}
+This paper suggests a way to circumvent these trade-offs and presents a generalization of the two LoRA fine-tuning methods for multi-task problems, which we call Hierarchical Low-Rank Adaption method (HLoRA).
+The Hierarchical LoRA method allows tasks to share information between each other through sharing a set of global hierarchical prior parameters. 
+This means that tasks with little data can benefit from the global structure of the related tasks, while tasks with more data can specialize on their specific task.
+We show in Section \ref{sec:hlora} that HLoRA can be seen as a generalization of the two approaches above.
+
+To test our method, we evaluate the Hierarchical LoRA method on the next token generation problem of parliament speeches letting each parliament representative be a task.
+In our setting, HLoRA is superior to both conventional approaches above and suggest a way to mitigate the trade-off between the two.
+% Write something about how easy it is to implement and how it can be used in practice?
+
+The paper is organized as follows: Section \ref{sec:related_works} presents related works, Section \ref{sec:method} presents the Hierarchical LoRA method, (...)
+
+% Should this fact be added somewhere?
+% If the D tasks are more similar to eachother than the pretraining corpus, the optimal parameters $\hat{\theta}_d$ should share some information during training in order to maximize their likelihood.
+
+\section{Related works} \label{sec:related_works}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\section{Method: Hierarchical LLM}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Method: Hierarchical LLM} \label{sec:method}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 This section describes the Hierarchical LoRA method for multi-task problems.
@@
 
-\subsection{The Hierarchical LoRA model}
+\subsection{The Hierarchical LoRA model} \label{sec:hlora}
 In the hierarchical LoRA model, we introduce a LoRA parameter set for each task $d$, denoted as $\theta_d$, containing all the low-rank decomposition matricies $A$ and $B$ from each decomposed layer in the Large Language Model.
@@
 %
+%
+The precision hyperparameter $\tau$ controls how much the task parameters $\theta_d$ should be allowed to deviate from the hierarchical mean parameters $\Theta$. 
+This is essentially a proxy for how much structure and information the tasks share with each other, and should be set by the practitioner based on her's prior knowledge.
+
+The resulting HLoRA model balances each task based on its data size and similarity to the other tasks: If a task has few data points, the posterior in equation \ref{eq:posterior} will be dominated by the prior term and the task parameters will be close to the global hierarchical parameters and borrow information and structure from these.
+On the other hand, if a task has a lot of data points, the likelihood will dominate the posterior and the task parameters will be allowed to learn more about the specific task and rely less on the hierarchical parameters.
+
+The hierarchical LoRA model can be seen as a generalization of the two conventional approaches for multi-task problems:
+If we let $\tau$ go to zero, the probability distributyion of the prior (eq.
+\ref{eq:prior}) will become a flat uniform prior and the problem will be equivalent to training each task independently.
+On the other hand, if we let $\tau$ go to infinity, the prior will become a strong constraint on the task parameters not allowing them to deviate from the global parameters, and the problem will be equivalent to training one model for all tasks.
 
@@
 \section{Experimental setup}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Use the Talk of Norway Dataset consisting of speeches of Norwegian parliament politicians.
 We will consider different speakers of this dataset as different tasks, and see how the hierarchical model can benefit from sharing information between these tasks.
-
+% add figure of the dataset distribution in file training_lengths.png
+\begin{figure}[h] \label{fig:task_len}
+    \centering
+    \includegraphics[width=\textwidth]{figures/training_lengths.png}
+    \caption{Number of training examples for each task}
+    \label{fig:training_lengths}
+\end{figure}
 % As base model, use opt-350m as it is predominantly trained on English, and only has seen a small amount of Norwegian data from the commoncrawls dataset.
 To test our method, we optimize the model with a range of different precision hyperparameters $\tau$, and evaluate the model's perplexity on the test set for each task.
-
 We use the Talk of Norway Dataset \cite{lapponi_talk_2018}, which consists of speeches of Norwegian parliament politicians.
@@
 
-% add figure of the dataset distribution in file training_lengths.png
-\begin{figure}[h] \label{fig:task_len}
-    \centering
-    \includegraphics[width=\textwidth]{figures/training_lengths.png}
-    \caption{Number of training examples for each task}
-    \label{fig:training_lengths}
-\end{figure}
-
 We select a model that has been pretrained on a large and diverse dataset, but has not seen the specific tasks that we want to evaluate it on.
@@
 
-Adapting learning rate to the precision hyperparameter..
+\section{Results}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The overall results of the hierarchical LoRA model on the Talk of Norway Dataset are shown in Table \ref{tbl:results} and Figure \ref{fig:results_plot}.
+We see that the model achieves the best test perplexity of 12.82 when using a precision parameter of $100$.
+The improvement is largest from training the models independently ($\tau = 0$), and smaller when comparing to the limiting case when all task specific model parameters are constraint to be equal ($\tau=10000$).
 
-\section{Results}
-The results of the hierarchical LoRA model on the Talk of Norway Dataset are shown in Table \ref{tbl:results} and Figure \ref{fig:results_plot}.
-We see that the model achieves the best validation perplexity of 12.82 when using a learning rate of $10^{-3}$ and a regularization constant of $10^{2}$.
-This indicates that the model benefits from sharing information between the tasks, and that the global prior helps to improve the performance of the individual task parameters.
-
-\begin{figure}[h]
+\begin{table}[h] \label{tbl:results}
     \centering
-    \includegraphics[width=\textwidth]{figures/results_plot.png}
-    \caption{Plot of Regularization Constant vs Perplexity}
-    \label{fig:results_plot}
-\end{figure}
-
-\begin{table}[h]
-\centering
-\caption{Empirical results of the hierarchical LORA finetuned on the three tasks. The perplexity is calculated on the test set.}
-\label{tbl:results}
     \begin{tabular}{ccc}
         \toprule
-    Learning Rate & Regularization & Validation Perplexity \\
+        Learning Rate & Precision ($\tau$) & Test Perplexity \\
         \midrule
@@
         \end{tabular}
+        \caption{Empirical results of the hierarchical LORA finetuned on the three tasks. The perplexity is calculated on the test set.}
     \end{table}
 
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=\textwidth]{figures/results_plot.png}
+    \caption{Plot of Precision vs Perplexity. The thick black line is the overall test perplexity over all tasks, and the thinner grey lines are the test perplexity for each individual task.}
+    \label{fig:results_plot}
+\end{figure}
+
+If we consider each task separately (Table \ref{tbl:task_perplexities}), we see that all tasks benefit from a hierarchical model compared to both the case when all models are trained independently ($\tau = 0$) and the limiting case when all task specific model parameters are constraint to be equal ($\tau=10000$).
+Figure \ref{fig:relative_improvement} shows the relative improvement of each task when using the hierarchical model compared to the two cases versus their training data size.
+There is an indication that the improvement from training separately is larger for tasks with less data, and we do not see this for the "one model" case.
+A precision parameter of $100$ gives the best perplexity for all our tasks.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{figures/relative_improvement_combined.png}
+    \caption{Task dataset size versus the relative improvement in perplexity of the best performing hierarchical model ($\tau = 100$) compared to the case when all models are trained independently ($\tau = 0$) in blue, and compared to the limiting case when all task specific model parameters are constraint to be equal ($\tau=10000$) in red.
+    The figure shows that all tasks benefit from sharing parameters with the global model, and indicates that the tasks with less data benefit more than the tasks with more data.}
+    \label{fig:relative_improvement}
+\end{figure}
+
 \section{Discussion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Our results in Figure \ref{fig:results_plot} show that the hierarchical LoRA model can improve the test perplexity for all tasks compared to both training the models independently and training one model for all tasks.
+This confirms our hypothesis that HLoRA can be a useful method for multi-task problems, and that it can mitigate the trade-off between training each task separately and training one model for all tasks.
+
+We observe in Figure \ref{fig:weights_vs_datalen} that the distance between the task parameters and the global prior increases with increasing task dataset size.
+This is as expected in a hierarchical model, as tasks with more data will have a larger likelihood term in the posterior and will be allowed to specialize more on their specific task during training.
+
+We find no relationship between dataset size and the final perplexity (Figure \ref{fig:absolute_improvement}).
+One could expect that tasks with more data are able to learn more about their specific task and achieve a lower perplexity, but this is not the case in our experiments.
+There are many other factors that can influence the final perplexity, and we believe that this is the reason in our case.
+However, when considering the relative improvement of the best performing model, the results in Figure \ref{fig:relative_improvement} show tasks with less data may benefit more than the tasks with more data.
+Although a weak signal, this is an expected result, as the hierarchical model allows tasks with little data to borrow information and structure from the global model, while tasks with more data can specialize on their specific task.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{figures/trainingsize_vs_perplexity.png}
+    \caption{Training size vs perplexity of the best performing hierarchical model ($\tau = 100$). The figure shows that there is no strong relationship between dataset size and final perplexity.}
+    \label{fig:absolute_improvement}
+\end{figure}
 
@@
 
-%\end{document}
+\section{Conclusion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+This paper presented a generalization of the LoRA fine-tuning method for multi-task problems, which we call the Hierarchical LoRA method.
+The hierarchical model allows tasks to share information between each other through a set of global hierarchical prior parameters, and can be seen as a generalization of the two conventional approaches for multi-task problems.
+We showed that the Hierarchical LoRA model can improve the test perplexity for all tasks compared to both training the models independently and training one model for all tasks, and that it can mitigate the trade-off between the two approaches.
 
-\section{Conclusion}
-
-\section*{Background}
-Low-Rank  Adaption (LoRA) of an existing large-scale pre-trained Large Language Models (LLM) \cite{hu_lora_2021} has become a popular go-to approach when solving a specific text problem.
-
-LoRA can reduce the number of trainable parameters by 10'000 times and the GPU memory requirement by 3 times \cite{hu_lora_2021}, while still outperform other fine-tuning techniques.
-We are interested in LoRA adaptions for multi-task problems, such as the TULU-v2-mix \cite{ivison_camels_2023}.
-
-For multi-task problems there are two conventional ways of training them. You can either 
-\begin{enumerate}
-    \item finetune each task separately
-    \item train one model for all tasks (given that the textual problem contains which task is being solved for each case)
-\end{enumerate}
-
-In addition to these two approaches, we present generalization of the two approaches: the Hierarchical Low-Rank Adaption.
-Given a pretrained LLM and $n$ tasks, we construct $n$ sets of LoRA parameters $\theta_n$ that each are trained on its separate task, like in approach (1) above.
-In additon, we create a global parameter set $\Theta$ that works as a prior to all the task specific parameters (or be regularized towards in a frequentist vocabulary).
-The final posterior will look something like this:
-%
-\begin{equation} \label{eq:posterior}
-    \sum_{n=1}^N  \sum_{i \in D_n} P(text_i | \theta_n) + \sum_{n=1}^N N(\theta_n | \Theta, \sigma)
-\end{equation}
-%
-where
-$D_n$ is the dataset corresponding to task $n$,
-$P(text_i | \theta)$ is the likelihood of sentence $text_i$ evaluated on the parameters $\theta$,
-
-$\sigma$ is a tuning parameter for how close the task specific parameters should be to the global parameter.
-
-\paragraph{Why Hierarchical LLM?}
-The Hierarchical LoRA Language Model will allow tasks to share information between each other through the global parameter set $\Theta$.
-If a task dataset has little data, the posterior  in equation (\ref{eq:posterior}) will mainly focus on the prior, whereas if the task dataset has a lot of data, the parameters can learn more about the specific task and move away from the global "average" model.
-
-The model above can be seen as a generalisation of the two approaches above.
-As $\sigma$ goes towards zero the problem will be equal to approach (2) (assuming we have added the task in the text), 
-and as $\sigma$ goes to infinity we get $n$ independent sets of parameters optimized independently as in approach (1).
-
-\paragraph{Research Questions}
-We have the following research questions
+Future work:
 \begin{itemize}
-    \item Will a maximum a posteriori hierarchical LoRA LLM  with one adapter per task outperform a single LoRA with the equivalent number of trainable parameters?
-    \item Privacy: Will training a Hierarchical LoRA on multiple tasks be more privacy friendly than training on two tasks jointly?
-    \item Will a posterior distribution of a LoRA LLM have "better" diversity/exploration than the sampling techniques currently used (p-sampling, nucleus samplng)
+    \item Investigate the effect of the hierarchical model on more tasks and datasets
+    \item Investigate the capacity of the global model
+    \item Investigate with varying capacity
+    \item Compute the full posterior and see if it can be used to explore better in text generation
 \end{itemize}
-
-\paragraph{Evaluation}
-We want to evaluate our research questions by (1) evaluating it on an open benchmark (TULU-v2-mix \cite{ivison_camels_2023}), (2) an evaluation by an GPT-4 agent, and (3) an online evaluation as a journalistic assistant in a news media article CMS, suggesting article attributes such as titles and tags.
 
@@
 
+\appendix
+\section{Task specific test perplexities}
+Table \ref{tbl:task_perplexities} shows the test perplexities for each task when using the hierarchical LoRA model with different precision hyperparameters $\tau$.
+
+\begin{table}[ht] \label{tbl:task_perplexities}
+    \centering
+    \caption{Test perplexity for each task and precision hyperparameter $\tau$}
+    \begin{tabular}{lrrrrrr}
+        \toprule
+        tau & 0     & 1     & 10    &          100   & 1000  & 10000 \\
+        Task                   &       &       &       &                &       &       \\
+        \midrule
+        abid q raja            & 15.56 & 15.20 & 12.57 & \textbf{11.60} & 11.91 & 12.41 \\
+        akhtar chaudhry        & 14.78 & 14.37 & 12.01 & \textbf{11.43} & 11.66 & 12.06 \\
+        aksel hagen            & 15.88 & 15.30 & 12.68 & \textbf{11.84} & 12.26 & 12.61 \\
+        alf egil holmelid      & 16.96 & 16.86 & 13.97 & \textbf{12.66} & 13.35 & 14.51 \\
+        anders b werp          & 16.86 & 16.35 & 13.04 & \textbf{11.87} & 12.54 & 13.00 \\
+        andré n skjelstad      & 15.53 & 15.23 & 12.59 & \textbf{11.58} & 11.98 & 12.41 \\
+        andré oktay dahl       & 16.01 & 15.85 & 13.50 & \textbf{12.55} & 12.88 & 13.48 \\
+        anette trettebergstuen & 15.98 & 15.73 & 13.43 & \textbf{12.57} & 12.90 & 13.28 \\
+        anita apelthun sæle    & 25.44 & 25.42 & 21.04 & \textbf{19.21} & 20.16 & 21.37 \\
+        anne tingelstad wøien  & 16.41 & 16.23 & 13.60 & \textbf{12.43} & 12.90 & 13.54 \\
+        anniken huitfeldt      & 14.44 & 14.39 & 12.33 & \textbf{11.66} & 12.05 & 12.53 \\
+        ansgar gabrielsen      & 16.10 & 16.16 & 14.07 & \textbf{13.23} & 13.68 & 14.19 \\
+        arild grande           & 16.65 & 16.29 & 13.31 & \textbf{12.21} & 12.56 & 13.08 \\
+        arne lyngstad          & 15.57 & 15.28 & 12.12 & \textbf{11.25} & 11.66 & 12.08 \\
+        arve kambe             & 17.78 & 17.40 & 14.24 & \textbf{13.11} & 13.43 & 13.79 \\
+        asmund kristoffersen   & 17.15 & 16.75 & 13.63 & \textbf{12.61} & 13.03 & 13.61 \\
+        audun lysbakken        & 15.21 & 15.25 & 13.22 & \textbf{12.48} & 12.67 & 13.13 \\
+        bendiks h arnesen      & 14.34 & 14.23 & 12.04 & \textbf{11.11} & 11.42 & 11.97 \\
+        bente thorsen          & 16.35 & 15.88 & 12.87 & \textbf{11.60} & 12.28 & 13.06 \\
+        bjørg tørresdal        & 16.02 & 15.65 & 12.78 & \textbf{11.58} & 11.92 & 12.46 \\
+        bjørn hernæs           & 16.33 & 15.94 & 13.68 & \textbf{12.76} & 13.08 & 13.58 \\
+        bjørn jacobsen         & 22.28 & 22.87 & 19.54 & \textbf{18.52} & 19.05 & 20.53 \\
+        bjørn lødemel          & 22.91 & 24.09 & 19.65 & \textbf{18.08} & 19.39 & 21.08 \\
+        bjørn tore godal       & 16.57 & 16.30 & 14.55 & \textbf{14.02} & 14.22 & 14.48 \\
+        borghild tenden        & 14.42 & 14.84 & 12.12 & \textbf{11.10} & 11.68 & 12.41 \\
+        \bottomrule
+    \end{tabular}
+\end{table}
+
 \end{document}

### 2024-03-31 src/generate_plots.py
update generate plot code

@@ -70,11 +70,17 @@ def get_perplexity_for_all_tasks(main_results):
 #%% ### ###
 ### Precompute values
 ### ### ###
-main_results = pd.read_csv("main_results.csv")
+main_results = pd.read_csv("figures/main_results.csv")
 df_task_perplexity, df = get_perplexity_for_all_tasks(main_results)
 dataset_sizes = build_dataset_of_dataset_sizes()
 tasks = df_task_perplexity['task'].unique()
-sns.set_theme(style='ticks', rc={'figure.figsize':(11.7,8.27)})
+
+cm=1/2.54
+sns.set_theme(
+    style='ticks', 
+    rc={'figure.figsize':(2*8.5*cm, 8.5*cm), 
+    'font.size': 8.5, 'axes.titlesize': 8, 'axes.labelsize': 8, 'xtick.labelsize': 8, 'ytick.labelsize': 8}
+    )
 
 #%% ### ###
 ### Generate latex table with main result
@@ -83,14 +89,58 @@ sns.set_theme(style='ticks', rc={'figure.figsize':(11.7,8.27)})
 df_main_results = pd.DataFrame(main_results)
 
 # Define formatters
+min_val_perplexity = df_main_results['validation perplexity'].min()
+
 formatters = {
-    'Learning Rate': lambda x: '$10^{{{}}}$'.format(int(np.log10(x))) if x != 0 else '0',
-    'Regularization': lambda x: '$10^{{{}}}$'.format(int(np.log10(x))) if x != 0 else '0',
-    'Validation Perplexity': '{:.2f}'.format
+    'lr': lambda x: '$10^{{{}}}$'.format(int(np.log10(x))) if x != 0 else '0',
+    'tau': lambda x: '$10^{{{}}}$'.format(int(np.log10(x))) if x != 0 else '0',
+    'validation perplexity': lambda x: '\\textbf{{{:.2f}}}'.format(x) if x == min_val_perplexity else '{:.2f}'.format(x)
 }
+# remove experiment_id
+df_main_results = df_main_results.drop(columns=['experiment_id'])
+
 
 # Print DataFrame as LaTeX table with formatters
 print(df_main_results.to_latex(index=False, formatters=formatters, escape=False))
+# SOME MANUAL ADJUSTMENT MADE TO TABLE!
+
+#%%
+task_specific_perplexity = df.copy()
+# drop columns: lr, validation perplexity, experiment_id
[…]

### 2024-05-01 paper/main.tex
Updates from Overleaf

@@
 
-\title{HLoRA: Hierarchical Low-Rank Adaption for Multi-task Large Language Models}
+\title{HLoRA: Hierarchical Bayesian Low-Rank Adaption for Multi-task Large Language Models}
 \author[1]{Simen Eide}
@@
 \maketitle
-\abstract{Lorem Ipsum}
+\abstract{Lorem Ipsum; also remember to write about the learning rate $\tau$}
 
 \section{Introduction}
-Large Language Models (LLM) have become a popular model architechture for solving a wide range of text problems.
-LLMs are typically pretrained on a large and diverse dataset, and then fine-tuned on a specific task.
+Large Language Models (LLM) have become a popular model for solving a wide range of generative text problems.
+LLMs are typically pretrained on a large and diverse dataset of texts, and then fine-tuned on a new data set of texts, dedicated to a specific task.
 The fine-tuning process is often done by maximizing the next token log likelihood of the task-specific dataset, and the resulting model is then used to make predictions and generations on new data.
-A popular fine-tuning method is Low-Rank Adaption (LoRA), which reduces the number of trainable parameters by up to 10'000 times and the GPU memory requirement by 3 times, while still outperforming other fine-tuning techniques (\cite{hu_lora_2021}).
+A popular fine-tuning method is Low-Rank Adaption (LoRA), which reduces the number of trainable parameters by a factor of up to 10'000 and the GPU memory requirement by a factor of 3, while still outperforming other fine-tuning techniques (\cite{hu_lora_2021}).
 
-When applying fine-tuning of LLMs to problems, the practitioner often wants to solve multiple similar tasks at the same time. 
-For instance, a news organization might want to use an LLM to generate various textual elements for a given article such as headlines, keywords, and summaries, all based on the main body of the text.
-This would involve fine-tuning the model for each of these distinct but related tasks.
-Similarly, a customer service chatbot might need to be fine-tuned to handle inquiries across a range of topics, such as product information, order status, and troubleshooting.
-Additionally, the chatbot might need to adapt its responses based on the user's tone or the formality of the conversation, requiring further fine-tuning.
-In a more creative context, a scriptwriting AI might be fine-tuned to generate dialogues for different characters in a movie, each with their unique style and vocabulary.
-This would involve fine-tuning the model for each character, a set of related but distinct tasks.
+When performing fine-tuning of LLMs, the practitioner often wants to solve multiple similar tasks at the same time: 
+A media organization might want to use an LLM to generate various textual elements to generate headlines, keywords, and summaries for a given news article. 
+A website might have multiple chatbot tasks answering questions to their users handling inquiries across a range of topics such as product information, order status, and troubleshooting.
+And finally, a scriptwriting AI might want to generate dialogues for different characters in a movie, each with their unique style and vocabulary.
 
-In these cases, the practitioner can either (1) train one model for each task separately, or (2) train one model for all tasks and condition the specific output wanted (e.g \cite{Raffel2019}).
-There is a trade-off between these two approaches: Train each task separately and the model will be able to specialize on each task, but it may suffer from limited data for each task.
+In these multi-task cases, the practitioner can either (1) train one model for each task and dataset separately, or (2) train one model for all tasks and condition the specific output wanted (e.g \cite{Raffel2019}).
+There is a trade-off between these two approaches: Train each task separately and the model will be able to specialize on each task, but it may suffer from limited data available for each task.
 Train one model for all tasks, and the model will be able to share information between tasks, but it may not be able to specialize on each task due to limited model capacity.
 
-This paper suggests a way to circumvent these trade-offs and presents a generalization of the two LoRA fine-tuning methods for multi-task problems, which we call Hierarchical Low-Rank Adaption method (HLoRA).
-The Hierarchical LoRA method allows tasks to share information between each other through sharing a set of global hierarchical prior parameters. 
+In this paper we suggest a way to circumvent these trade-offs and presents a generalization of the two LoRA fine-tuning methods for multi-task problems, which we call Hierarchical Low-Rank Adaption method (HLoRA).
+The Hierarchical LoRA method allows tasks to share information between each other through sharing a set of global hierarchical prior parameters, in a Bayesian hierarchical setting. 
 This means that tasks with little data can benefit from the global structure of the related tasks, while tasks with more data can specialize on their specific task.
@@
 Multi-task Learning \cite{caruana_multitask_nodate} is a popular machine learning approach for training one model on multiple tasks.
-A common approach is to share the lower layers of a neural network between tasks, and have separate output layers for each task.
-Our paper focuses on a different approach, where we share information between tasks through a hierarchical prior over the task-specific parameters.
+A common approach is to share the lower layers of a neural network between tasks, and have separate parameters for the last (or last few) layers for each task.
+Our paper suggests a different approach, where we share information between tasks through a hierarchical prior over the task-specific parameters.
 
@@
 This approach has been widely adopted in neural networks.
-However, we focus on the LoRA method, which allows us to train a low rank decomposition of the entire network.
+However, in this paper we focus on the LoRA method, which instead allows us to train a low rank decomposition of the entire network.
 
@@
 Another variation, the Weight Decomposed Low-Rank Adaption (DoRA) \cite{hayou_lora_2024}, decomposes the low-rank parameters into scale and unit norm direction parameters.
-There's also the Efficient Low Rank Adaption (LoRA+) \cite{hayou_lora_2024}, which improves the existing LoRA optimization algorithm by adjusting the learning rate of the low-rank parameters.
-All these methods are designed to improve the performance of the LoRA algorithm, and can be used in conjunction with the Hierarchical LoRA method.
+There is also the Efficient Low Rank Adaption (LoRA+) \cite{hayou_lora_2024}, which improves the existing LoRA optimization algorithm by adjusting the learning rate of the low-rank parameters.
+All these methods are designed to improve the performance of the LoRA algorithm, and can be used in conjunction with the proposed Hierarchical LoRA method.
 
@@
 We define our study across $D$ distinct tasks, denoted as $d=1,2,...,D$.
-For each task $d$, we have a dataset $\mathcal{D}_d$ containing $N_d$ documents, with each document consisting of a sequence of $W_n$ tokens.
+For each task $d$, we have a dataset $\mathcal{D}_d$ containing $N_d$ documents, with each document consisting of a sequence of $W_n$ tokens, $n=1,..., N_d$.
 This dataset structure can be formally represented as:
@@
 The model's probability distribution is given by:
-The model's probability distribution is given by:
 
@@
 where $LLM(w_{i+1} | w_{1:i})$ is the probability of the next token $w_{i+1}$ given the previous tokens $w_{1:i}$ in the document.
-We omit to write $\theta_{full}$ explicitly in the model, as the autoregressive language model always depend on these parameters.
+We omit to write $\theta_{full}$ explicitly in the LLM, as the autoregressive language model always depend on these parameters.
 
 \subsection{Low-Rank Adaption (LoRA)}
-A Large Language Model contains many linear layers, and the LoRA method reduces the number of trainable parameters by reparameterizing the weight parameters ($W \in \R^{n_1 \times n_2}$) of these layers into the original pretrained weights $W_{full}$ and a low-rank decomposition $A$ and $B$ (\cite{hayou_lora_2024}):
+A Large Language Model contains many linear layers, and the LoRA method reduces the number of trainable parameters by reparameterizing the weight parameters ($W \in \R^{n_1 \times n_2}$, $W \subset \theta$) of these layers into the original pretrained weights $W_{full}$ and a low-rank decomposition $A$ and $B$ (\cite{hayou_lora_2024}):
 \begin{equation} \label{eq:LoRA}
@@
 %
-where $W_{full} \in \R^{n_1 \times n_2}$ is the original pretrained weight, $B \in \R^{n_1 \times r}$ and $A \in \R^{r \times n_2}$ are the low-rank factors, $r \ll \text{min}(n_1, n_2)$ is the low rank dimension and $\alpha$ is a scaling hyperparameter.
+where 
+$W_{full} \in \R^{n_1 \times n_2}$ is the original pretrained weight, 
+$B \in \R^{n_1 \times r}$ and $A \in \R^{r \times n_2}$ are the low-rank factors, 
+$n_1$, $n_2$ are the input and output dimension of the original weight matrix in the linear layer,
+$r \ll \text{min}(n_1, n_2)$ is the low rank dimension 
+and $\alpha$ is a scaling hyperparameter.
 
-During LoRA fine-tuning, the low-rank factors $A$ and $B$ are learned from data as normal parameter matricies, while the original pretrained weights $W_{full}$ are kept fixed.
+During LoRA fine-tuning, the low-rank factors $A$ and $B$ are learned (estimated) from data, while the original pretrained weights $W_{full}$ are kept fixed to their pretrained value.
 
 \subsection{The Hierarchical LoRA model} \label{sec:hlora}
-In the hierarchical LoRA model, we introduce a LoRA parameter set for each task $d$, denoted as $\theta_d$, containing all the low-rank decomposition matricies $A$ and $B$ from each decomposed layer in the Large Language Model.
+In the hierarchical LoRA model, we introduce a LoRA parameter set for each task $d$, denoted as $\theta_d$, containing all the low-rank decomposition matricies $A_d$ and $B_d$ from each decomposed layer in the Large Language Model.
 The likelihood of the data $\mathcal{D}$ given the task parameters $\theta_{1:D}$ is then given by (combining equation \ref{eq:data} and \ref{eq:LLMprob}):
 \begin{equation} \label{eq:lik}
-    \text{L}(\mathcal{D} | \theta_1, \theta_2, \ldots, \theta_D) = \prod_{d=1}^D L(\mathcal{D}_d | \theta_d) = \prod_{d=1}^D \prod_{n=1}^{N_d} \prod_{i=1}^{W_n-1} \text{LLM}(w_{d,n,i+1} | w_{d,n,1:i}; \theta_d)
+    \text{L}(\mathcal{D} | \theta_1, \theta_2, \ldots, \theta_D) = \prod_{d=1}^D L(\mathcal{D}_d | \theta_d) = \prod_{d=1}^D \prod_{n=1}^{N_d} \prod_{i=1}^{W_n-1} \text{LLM}(w_{d,n,i+1} | w_{d,n,1:i}; \theta_d).
 \end{equation}
@@
 %
-Each task will share some information with each other, and we model this by introducing a Gaussian hierarchical prior over the task parameters $\theta_{1:D}$, given by
+Tasks will share some knowledge with each other, and we model this by introducing a Gaussian hierarchical prior over the task parameters $\theta_{1:D}$, given by
 \begin{equation} \label{eq:prior}
@@
 %
-where $\Theta$ is a set of hierarchical mean parameters, $I$ is a unit diagonal matrix and $\tau \geq 0$ is a scalar precision hyperparameter controlling how similar task parameters $\theta_d$ should be to the hierarchical mean parameters $\Theta$.
-We let $\Theta$ have a uniform prior (i.e. $P(\Theta) = 1$).
+where $\Theta$ is a set of hierarchical mean parameters, $I$ is a unit diagonal matrix and $\tau \geq 0$ is a scalar precision hyperparameter controlling how similar the task parameters $\theta_d$ should be to the hierarchical mean parameters $\Theta$.
+We let $\Theta$ have a uniform hyperprior (i.e.
+$P(\Theta) = 1$).
+The prior assumes that the task parameters share the common origin from the shared Gaussian prior, because the tasks are related: input data come from similar distributions and the output tasks has similar aspects.
+The precision hyperparameter $\tau$ controls how much the task parameters $\theta_d$ are a priori expected to be allowed to deviate from the hierarchical mean parameters $\Theta$. 
+This is essentially a measure for how much structure and information the tasks share with each other, and should be set by the practitioner based on her's prior knowledge.
 
@@
 %
-The precision hyperparameter $\tau$ controls how much the task parameters $\theta_d$ should be allowed to deviate from the hierarchical mean parameters $\Theta$. 
-This is essentially a proxy for how much structure and information the tasks share with each other, and should be set by the practitioner based on her's prior knowledge.
 
@@
 The hierarchical LoRA model can be seen as a generalization of the two conventional approaches for multi-task problems:
-If we let $\tau$ go to zero, the probability distributyion of the prior (eq.
+If we let $\tau$ go to zero, the probability distribution of the prior (eq.
 \ref{eq:prior}) will become a flat uniform prior and the problem will be equivalent to training each task independently.
@@
 \subsection{Optimization}
-Given a precision hyperparameter, we want to find the maximum a posteriori (MAP) estimate of the hierarchical model. 
-We will do this by using AdamW (\cite{adamW}), a gradient-based optimization algorithm, optimizing over the task parameters $\theta_{1:D}$ and the hierarchical mean parameters $\Theta$ simultaneously optimizing the likelihood and the prior in equation \ref{eq:posterior}.
-This implies that the task parameters with have gradients both towards minimizing the convential LLM loss and the hierarchical mean parameters.
-The hierarchical mean parameters $\Theta$ will have gradients towards the average of the task parameters, and learn to represent the global structure of the tasks after the task parameters have been partly optimized.
+Given a precision hyperparameter $\tau$, we want to find the maximum a posteriori (MAP) estimate of the hierarchical model. 
+We will do this by using AdamW (\cite{adamW}), a gradient-based optimization algorithm, optimizing over the task parameters $\theta_{1:D}$ and the hierarchical mean parameters $\Theta$ using the posterior in equation \ref{eq:posterior}.
+This implies that the task parameters with have gradients both towards minimizing the convential LLM loss (represented by the likelihood term) and the hierarchical mean parameters (originating from the prior).
+The hierarchical mean parameters $\Theta$ will have gradients towards the average of the task parameters, and learn to represent the global structure of the tasks.
 
 An important observation is that the gradient of the prior term in equation \ref{eq:posterior} is proportional to the precision hyperparameter $\tau$.
-This means that the optimization will more steps to converge for larger values of $\tau$, as the task parameters will be more constrained towards the hierarchical mean parameters.
-Therefore, we adjust the learning rate of the AdamW optimizer to be proportional to the precision hyperparameter $\tau$, and we find in informal tests that this cause the optimizations to converge at a similar rate for different values of $\tau$.
+This means that the optimization will need more steps to converge for larger values of $\tau$, as the task parameters will be more constrained towards the hierarchical mean parameters, and a posteriori dependent on each other.
+Therefore, we adjust the learning rate of the AdamW optimizer to be proportional to the precision hyperparameter $\tau$, and we find in our experiments that this causes the optimizations to converge at a similar rate for different values of $\tau$.
 
@@
 % As base model, use opt-350m as it is predominantly trained on English, and only has seen a small amount of Norwegian data from the commoncrawls dataset.
-To test our method, we optimize the model with a range of different precision hyperparameters $\tau$, and evaluate the model's perplexity on the test set for each task.
+To test our method, we optimize the model with a range of different precision hyperparameters $\tau$, and evaluate the models' perplexity on the test set for each task.
 We use the Talk of Norway Dataset \cite{lapponi_talk_2018}, which consists of speeches of Norwegian parliament politicians.
@@
 Specifically, we select chronologically the 25 first speakers that have above 100 speeches in the dataset.
-This gives us a total of 25 tasks with samples ranging from 110 to 477 documents, and an average of 202 documents per task.
+This gives us a total of 25 tasks with samples ranging from 110 to 477 documents, and an average of 202 documents per task. For some speakers we have significantly less data as for others. 
 The distibution of samples per task can be seen in Figure \ref{fig:task_len}.
@@
 
-\begin{table}[h] \label{tbl:results}
+\begin{table}[h] 
     \centering
@@
         \caption{Empirical results of the hierarchical LORA finetuned on the three tasks. The perplexity is calculated on the test set.}
+        \label{tbl:results}
     \end{table}
@@
 One could expect that tasks with more data are able to learn more about their specific task and achieve a lower perplexity, but this is not the case in our experiments.
-There are many other factors that can influence the final perplexity, and we believe that this is the reason in our case.
+There are many other task specific factors such as under-represented dialects and political views that can influence the final perplexity.
+Controlling for these factors is difficult and out of scope for this paper, but we believe that these factors may be the reason for this negative result.
+
 However, when considering the relative improvement of the best performing model, the results in Figure \ref{fig:relative_improvement} show tasks with less data may benefit more than the tasks with more data.
 Although a weak signal, this is an expected result, as the hierarchical model allows tasks with little data to borrow information and structure from the global model, while tasks with more data can specialize on their specific task.
+
+It is in principle possible to avoid the MAP maximisation, and instead take a fully Bayesian approach, which would need to rely on some form of MCMC.
+However, MCMC methods
+and in this sense might not scale.
+One advantage is the possibility to estimate $\tau$, and integrate over its empirical posterior distribution obtained within the MCMC.
+We have studied the effect of $tau$ in the test data.
+One can also cross-validate $tau$ in the training data.
+This would also be computationally too heavy.
+However, one could devise a way of subsampling the domain training data; for example, one could use only a few domains with large training data and few with little data, and run a rapid cross-validation for $tau$, the result of which would be used further.
+
+Another important discovery is the usefulness to let $tau$ determine the learning rate. The intiuition is clear and the effect potent.
 
@@
     \item Compute the full posterior and see if it can be used to explore better in text generation
+    \item Prove that the optimal learning rate is indeed determined by $\tau$. 
 \end{itemize}

### 2024-06-06 paper/main.tex
add abstract, conclusion and perplexity def

@@
 \maketitle
-\abstract{Lorem Ipsum; also remember to write about the learning rate $\tau$}
+\abstract{This paper introduces Bayesian Hierarchical Low-Rank Adaption (BH-LoRA), a novel method for finetuning multi-task Large Language Models (LLMs).
+Current finetuning approaches, such as Low-Rank Adaption (LoRA), perform exeptionally well in reducing training parameters and memory usage but face limitations when applied to multiple similar tasks.
+Practitioners usually have to choose between training separate models for each task or a single model for all tasks, both of which come with trade-offs in specialization and data utilization.
+
+HLoRA addresses these trade-offs by leveraging a Bayesian hierarchical model that allows tasks to share information through global hierarchical prior parameters.
+This enables tasks with limited data to benefit from the overall structure derived from related tasks while allowing tasks with more data to specialize.
+We demonstrate the effectiveness of HLoRA on the next token generation task using the Talk of Norway Dataset, where each parliament representative's speeches represent distinct tasks.
+Our experimental results show that HLoRA outperforms both traditional individual and unified model approaches, achieving lower perplexity and better generalization across tasks.
+This method provides a scalable and efficient solution for multi-task LLM finetuning, with significant practical implications for diverse applications such as media content generation, chatbot responses, and scriptwriting.
+}
 
@@
 The fine-tuning process is often done by maximizing the next token log likelihood of the task-specific dataset, and the resulting model is then used to make predictions and generations on new data.
-A popular fine-tuning method is Low-Rank Adaption (LoRA), which reduces the number of trainable parameters by a factor of up to 10'000 and the GPU memory requirement by a factor of 3, while still outperforming other fine-tuning techniques (\cite{hu_lora_2021}).
+A popular fine-tuning method is Low-Rank Adaption (LoRA), which reduces the number of trainable parameters by a factor of up to 10'000 and the GPU memory requirement by a factor of 3, while still outperforming other finetuning techniques (\cite{hu_lora_2021}).
 
@@
 % As base model, use opt-350m as it is predominantly trained on English, and only has seen a small amount of Norwegian data from the commoncrawls dataset.
-To test our method, we optimize the model with a range of different precision hyperparameters $\tau$, and evaluate the models' perplexity on the test set for each task.
-We use the Talk of Norway Dataset \cite{lapponi_talk_2018}, which consists of speeches of Norwegian parliament politicians.
+To test our method, we optimize the model with a range of different precision hyperparameters $\tau$, and evaluate the models' Perplexity on the test set for each task.
+% DEF PERPLEXITY
+Perplexity is defined as the exponentiated average negative log-likelihood per token, and is a common metric for evaluating language models. A lower perplexity indicates a better model.
+We use the Talk of Norway Dataset~\cite{lapponi_talk_2018}, which consists of speeches of Norwegian parliament politicians.
 Each speaker will come from different political parties and geographical areas, but the tasks will still share a common domain. 
-Therefore, we can consider different speakers of this dataset as different tasks.
-Specifically, we select chronologically the 25 first speakers that have above 100 speeches in the dataset.
+Therefore, we can consider different speakers of this dataset as different tasks. Specifically, we select chronologically the 25 first speakers that have above 100 speeches in the dataset.
 This gives us a total of 25 tasks with samples ranging from 110 to 477 documents, and an average of 202 documents per task. For some speakers we have significantly less data as for others. 
-The distibution of samples per task can be seen in Figure \ref{fig:task_len}.
+The distibution of samples per task can be seen in Figure~\ref{fig:task_len}.
 We leave 33\% of the data for the test set.
@@
 We select a model that has been pretrained on a large and diverse dataset, but has not seen the specific tasks that we want to evaluate it on.
-We will use the `opt-350m` model \cite{zhang_opt_2022}, which is predominantly trained on English, and has only seen a small amount of Norwegian data from the CommonCrawls dataset.
+We will use the `opt-350m' model~\cite{zhang_opt_2022}, which is predominantly trained on English, and has only seen a small amount of Norwegian data from the CommonCrawls dataset.
 
@@
 This confirms our hypothesis that HLoRA can be a useful method for multi-task problems, and that it can mitigate the trade-off between training each task separately and training one model for all tasks.
+It is interesting to note the HLoRA also helps for the tasks with relatively more data.
 
@@
 
-It is in principle possible to avoid the MAP maximisation, and instead take a fully Bayesian approach, which would need to rely on some form of MCMC.
-However, MCMC methods
-and in this sense might not scale.
-One advantage is the possibility to estimate $\tau$, and integrate over its empirical posterior distribution obtained within the MCMC.
-We have studied the effect of $tau$ in the test data.
-One can also cross-validate $tau$ in the training data.
-This would also be computationally too heavy.
-However, one could devise a way of subsampling the domain training data; for example, one could use only a few domains with large training data and few with little data, and run a rapid cross-validation for $tau$, the result of which would be used further.
-
-Another important discovery is the usefulness to let $tau$ determine the learning rate. The intiuition is clear and the effect potent.
+%% MCMC vs MAP
+In principle, one could bypass the Maximum a Posteriori maximization and adopt a fully Bayesian approach which would tell us how much we can trust the hierarchical model.
+Such an approach  would require the use of Markov Chain Monte Carlo, Variational Inference or other methods.
+This would allow us to sample from the posterior distribution of the hierarchical model, and obtain uncertainty estimates for the task parameters.
+However, MCMC methods are computationally heavy, and it is not clear to us how to best sample from the posterior distribution of the language model in an efficient and scalable way.
+A middle ground between a full bayesian approach and the MAP estimate is to be bayesian in a few parameters.
+For instance, instead of our approach where we maximize the likelihood of the validation set with respect to the precision parameter $\tau$, one could be bayesian and integrate over its empirical posterior distribution obtained within the MCMC.
+This would give us an estimate of how well the tasks share information with each other in the specific problem.
+We leave these questions for future work.
 
@@
 The hierarchical model allows tasks to share information between each other through a set of global hierarchical prior parameters, and can be seen as a generalization of the two conventional approaches for multi-task problems.
-We showed that the Hierarchical LoRA model can improve the test perplexity for all tasks compared to both training the models independently and training one model for all tasks, and that it can mitigate the trade-off between the two approaches.
-
-Future work:
-\begin{itemize}
-    \item Investigate the effect of the hierarchical model on more tasks and datasets
-    \item Investigate the capacity of the global model
-    \item Investigate with varying capacity
-    \item Compute the full posterior and see if it can be used to explore better in text generation
-    \item Prove that the optimal learning rate is indeed determined by $\tau$. 
-\end{itemize}
+We show that the Hierarchical LoRA model can improve the test perplexity for all tasks compared to both training the models independently and training one model for all tasks, and that it can mitigate the trade-off between the two approaches.
+Future work includes investigating the effect that HLORA has on more tasks and datasets, investigating how the relative capacity of the global model affects the performance, and investigating the full posterior distribution of the model.
 

### 2024-06-06 paper/main.tex
rename hlora to bora

@@
 
-\usepackage[
-backend=biber,
-style=alphabetic,
-sorting=ynt
-]{biblatex}
+\usepackage[backend=biber,style=alphabetic,sorting=ynt]{biblatex}
+\addbibresource{My Library.bib}
 \usepackage{amsmath}
@@
 \maketitle
-\abstract{This paper introduces Bayesian Hierarchical Low-Rank Adaption (BH-LoRA), a novel method for finetuning multi-task Large Language Models (LLMs).
+\abstract{This paper introduces Bayesian Hierarchical Low-Rank Adaption (BoRA), a novel method for finetuning multi-task Large Language Models (LLMs).
 Current finetuning approaches, such as Low-Rank Adaption (LoRA), perform exeptionally well in reducing training parameters and memory usage but face limitations when applied to multiple similar tasks.
@@
 
-HLoRA addresses these trade-offs by leveraging a Bayesian hierarchical model that allows tasks to share information through global hierarchical prior parameters.
+BoRA addresses these trade-offs by leveraging a Bayesian hierarchical model that allows tasks to share information through global hierarchical prior parameters.
 This enables tasks with limited data to benefit from the overall structure derived from related tasks while allowing tasks with more data to specialize.
-We demonstrate the effectiveness of HLoRA on the next token generation task using the Talk of Norway Dataset, where each parliament representative's speeches represent distinct tasks.
-Our experimental results show that HLoRA outperforms both traditional individual and unified model approaches, achieving lower perplexity and better generalization across tasks.
+We demonstrate the effectiveness of BoRA on the next token generation task using the Talk of Norway Dataset, where each parliament representative's speeches represent distinct tasks.
+Our experimental results show that BoRA outperforms both traditional individual and unified model approaches, achieving lower perplexity and better generalization across tasks.
 This method provides a scalable and efficient solution for multi-task LLM finetuning, with significant practical implications for diverse applications such as media content generation, chatbot responses, and scriptwriting.
@@
 
-In this paper we suggest a way to circumvent these trade-offs and presents a generalization of the two LoRA fine-tuning methods for multi-task problems, which we call Hierarchical Low-Rank Adaption method (HLoRA).
+In this paper we suggest a way to circumvent these trade-offs and presents a generalization of the two LoRA fine-tuning methods for multi-task problems, which we call Hierarchical Bayesian Low-Rank Adaption method (BoRA).
 The Hierarchical LoRA method allows tasks to share information between each other through sharing a set of global hierarchical prior parameters, in a Bayesian hierarchical setting. 
 This means that tasks with little data can benefit from the global structure of the related tasks, while tasks with more data can specialize on their specific task.
-We show in Section \ref{sec:hlora} that HLoRA can be seen as a generalization of the two approaches above.
+We show in Section \ref{sec:BoRA} that BoRA can be seen as a generalization of the two approaches above.
 
 To test our method, we evaluate the Hierarchical LoRA method on the next token generation problem of parliament speeches letting each parliament representative be a task.
-In our setting, HLoRA is superior to both conventional approaches above and suggest a way to mitigate the trade-off between the two.
+In our setting, BoRA is superior to both conventional approaches above and suggest a way to mitigate the trade-off between the two.
 % Write something about how easy it is to implement and how it can be used in practice?
 
-The paper is organized as follows: Section \ref{sec:related_works} presents related works, Section \ref{sec:method} presents the Hierarchical LoRA method, (...)
+The paper is organized as follows: Section \ref{sec:related_works} presents related works, Section \ref{sec:method} presents the Hierarchical LoRA method, Section \ref{sec:experimental_setup} describes the experimental setup, Section \ref{sec:results} presents the results, and Section \ref{sec:discussion} discusses the results and the implications of the Hierarchical LoRA method.
 
@@
 
-\subsection{The Hierarchical LoRA model} \label{sec:hlora}
+\subsection{The Hierarchical Bayesian LoRA model} \label{sec:BoRA}
 In the hierarchical LoRA model, we introduce a LoRA parameter set for each task $d$, denoted as $\theta_d$, containing all the low-rank decomposition matricies $A_d$ and $B_d$ from each decomposed layer in the Large Language Model.
@@
 
-The resulting HLoRA model balances each task based on its data size and similarity to the other tasks: If a task has few data points, the posterior in equation \ref{eq:posterior} will be dominated by the prior term and the task parameters will be close to the global hierarchical parameters and borrow information and structure from these.
+The resulting BoRA model balances each task based on its data size and similarity to the other tasks: If a task has few data points, the posterior in equation \ref{eq:posterior} will be dominated by the prior term and the task parameters will be close to the global hierarchical parameters and borrow information and structure from these.
 On the other hand, if a task has a lot of data points, the likelihood will dominate the posterior and the task parameters will be allowed to learn more about the specific task and rely less on the hierarchical parameters.
@@
 Our results in Figure \ref{fig:results_plot} show that the hierarchical LoRA model can improve the test perplexity for all tasks compared to both training the models independently and training one model for all tasks.
-This confirms our hypothesis that HLoRA can be a useful method for multi-task problems, and that it can mitigate the trade-off between training each task separately and training one model for all tasks.
-It is interesting to note the HLoRA also helps for the tasks with relatively more data.
+This confirms our hypothesis that BoRA can be a useful method for multi-task problems, and that it can mitigate the trade-off between training each task separately and training one model for all tasks.
+It is interesting to note the BoRA also helps for the tasks with relatively more data.
 
@@
 We show that the Hierarchical LoRA model can improve the test perplexity for all tasks compared to both training the models independently and training one model for all tasks, and that it can mitigate the trade-off between the two approaches.
-Future work includes investigating the effect that HLORA has on more tasks and datasets, investigating how the relative capacity of the global model affects the performance, and investigating the full posterior distribution of the model.
+Future work includes investigating the effect that BoRA has on more tasks and datasets, investigating how the relative capacity of the global model affects the performance, and investigating the full posterior distribution of the model.
 

### 2024-06-06 paper/main.tex
resize figures, tables and equations to fit

@@
 The likelihood of the data $\mathcal{D}$ given the task parameters $\theta_{1:D}$ is then given by (combining equation \ref{eq:data} and \ref{eq:LLMprob}):
-\begin{equation} \label{eq:lik}
-    \text{L}(\mathcal{D} | \theta_1, \theta_2, \ldots, \theta_D) = \prod_{d=1}^D L(\mathcal{D}_d | \theta_d) = \prod_{d=1}^D \prod_{n=1}^{N_d} \prod_{i=1}^{W_n-1} \text{LLM}(w_{d,n,i+1} | w_{d,n,1:i}; \theta_d).
-\end{equation}
+
+\begin{align}  \label{eq:lik}
+    & \text{L}(\mathcal{D} | \theta_1, \theta_2, \ldots, \theta_D) = \prod_{d=1}^D L(\mathcal{D}_d | \theta_d) \nonumber \\
+    &= \prod_{d=1}^D \prod_{n=1}^{N_d} \prod_{i=1}^{W_n-1} \text{LLM}(w_{d,n,i+1} | w_{d,n,1:i}; \theta_d).
+\end{align}
 %
@@
 % add figure of the dataset distribution in file training_lengths.png
-\begin{figure}[h] \label{fig:task_len}
+\begin{figure} \label{fig:task_len}
     \centering
+    \resizebox{\linewidth}{!}{%
     \includegraphics[width=\textwidth]{figures/training_lengths.png}
+    }
     \caption{Number of training examples for each task}
@@
 
+\begin{lstlisting}[gobble=2]
     \begin{table}[h] 
         \centering
+        \resizebox{\linewidth}{!}{%
             \begin{tabular}{ccc}
@@
                 \end{tabular}
+        }
             \caption{Empirical results of the hierarchical LORA finetuned on the three tasks. The perplexity is calculated on the test set.}
@@
         \end{table}
+\end{lstlisting}
 
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=\textwidth]{figures/results_plot.png}
+\begin{figure}%[ht]
+    %\centering
+    \resizebox{\linewidth}{!}{%
+    \includegraphics{figures/results_plot.png}
+    }
     \caption{Plot of Precision vs Perplexity. The thick black line is the overall test perplexity over all tasks, and the thinner grey lines are the test perplexity for each individual task.}
@@
     \centering
+    \resizebox{\linewidth}{!}{%
     \includegraphics[width=\textwidth]{figures/relative_improvement_combined.png}
+    }
     \caption{Task dataset size versus the relative improvement in perplexity of the best performing hierarchical model ($\tau = 100$) compared to the case when all models are trained independently ($\tau = 0$) in blue, and compared to the limiting case when all task specific model parameters are constraint to be equal ($\tau=10000$) in red.
@@
     \centering
+    \resizebox{\linewidth}{!}{%
     \includegraphics[width=\textwidth]{figures/trainingsize_vs_perplexity.png}
+    }
     \caption{Training size vs perplexity of the best performing hierarchical model ($\tau = 100$). The figure shows that there is no strong relationship between dataset size and final perplexity.}
@@
     \centering
+    \resizebox{\linewidth}{!}{%
     \includegraphics[width=\textwidth]{figures/weights_vs_data_lengths.png}
+    }
     \caption{The figure shows the L2-distance between each task's adapter weight and the global prior on the y-axis, and the number of training examples on the x-axis.
@@
     \caption{Test perplexity for each task and precision hyperparameter $\tau$}
+    \resizebox{\linewidth}{!}{%
     \begin{tabular}{lrrrrrr}
@@
     \end{tabular}
+    }
 \end{table}

### 2024-06-06 src/pca_analysis.py
update plotting

@@ -30,7 +30,7 @@ params = {
 tokenizer = load_tokenizer(params)
 dataloaders, task_stats = data_utils.prepare_talkofnorway_dataloaders(tokenizer, **params)
 
-pl_model = load_model(params, task_stats=task_stats, tokenizer=tokenizer, checkpoint_path="/home/simen.eide@schibsted.com/hier-llm/src/logs_regularized/1kepoch-reg:10-lr:0.0001-global:False/version_1/checkpoints/epoch=281-step=22560.ckpt")
+pl_model = load_model(params, task_stats=task_stats, tokenizer=tokenizer, checkpoint_path="/home/simen.eide@schibsted.com/hier-llm/src/logs/1kepoch-reg:10-lr:0.0001-global:False/version_1/checkpoints/epoch=281-step=22560.ckpt")
 
 model = pl_model.model
 #%% concat all adapter weights from different tasks
@@ -70,7 +70,12 @@ import seaborn as sns
 import numpy as np
 import matplotlib.pyplot as plt
 import seaborn as sns
-sns.set_theme()
+cm=1/2.54
+sns.set_theme(
+    style='ticks', 
+    rc={'figure.figsize':(2*8.5*cm, 8.5*cm), 
+    'font.size': 8.5, 'axes.titlesize': 8, 'axes.labelsize': 8, 'xtick.labelsize': 8, 'ytick.labelsize': 8}
+    )
 # Create a new figure and set its size
 fig, ax = plt.subplots(figsize=(10, 8))
 # Create a scatter plot
@@ -88,8 +93,8 @@ ax.tick_params(axis='both', which='major', labelsize=10)
 plt.grid(True)
 
 # Remove the top and right spines
-ax.spines['right'].set_visible(False)
-ax.spines['top'].set_visible(False)
+#ax.spines['right'].set_visible(False)
+#ax.spines['top'].set_visible(False)
 
 # Save the plot with a transparent background
 plt.savefig('weights_vs_data_lengths.png', bbox_inches='tight', dpi=300, transparent=True)

### 2024-06-06 paper/main.tex
nldl cls file
fix references

@@
-\documentclass{article}
+%\documentclass{article}
+\documentclass[fullpaper]{nldl}
+\paperID{42}
+\vol{V}
 \usepackage{graphicx} % Required for inserting images
@@
 
-\title{HLoRA: Hierarchical Bayesian Low-Rank Adaption for Multi-task Large Language Models}
+\title{BoRA: Hierarchical Bayesian Low-Rank Adaption for Multi-task Large Language Models}
 \author[1]{Simen Eide}
@@
 
-\usepackage[backend=biber,style=alphabetic,sorting=ynt]{biblatex}
-\addbibresource{My Library.bib}
+%\usepackage{biblatex} % [backend=biber,style=alphabetic,sorting=ynt] %not necessary when using template
+\addbibresource{references.bib}
 \usepackage{amsmath}
@@
 \usepackage{booktabs}
-\addbibresource{My Library.bib}
+%PACKAGES FROM TEMPLATE THAT SEEMS NECESSARY
+\usepackage{mathtools}
+\usepackage{graphicx}
+\usepackage{listings}
+\lstset{
+  basicstyle=\small\ttfamily,
+  breaklines,
+}
+\usepackage{enumitem}
+\usepackage{algorithm}
+\usepackage{algorithmic}
+\newcommand{\theHalgorithm}{\arabic{algorithm}}
+%%% END OF PACKAGES FROM TEMPLATE
 
@@
 The fine-tuning process is often done by maximizing the next token log likelihood of the task-specific dataset, and the resulting model is then used to make predictions and generations on new data.
-A popular fine-tuning method is Low-Rank Adaption (LoRA), which reduces the number of trainable parameters by a factor of up to 10'000 and the GPU memory requirement by a factor of 3, while still outperforming other finetuning techniques (\cite{hu_lora_2021}).
+A popular fine-tuning method is Low-Rank Adaption (LoRA), which reduces the number of trainable parameters by a factor of up to 10'000 and the GPU memory requirement by a factor of 3, while still outperforming other finetuning techniques (\cite{hu_lora_2022}).
 
@@
 \subsection{Multi-task Learning}
-Multi-task Learning \cite{caruana_multitask_nodate} is a popular machine learning approach for training one model on multiple tasks.
+Multi-task Learning \cite{caruana_multitask_1997} is a popular machine learning approach for training one model on multiple tasks.
 A common approach is to share the lower layers of a neural network between tasks, and have separate parameters for the last (or last few) layers for each task.
@@
 A popular approach to finetuning neural networks involves adjusting only the top layer (or the n top layers) of a pre-trained neural network (e.g.
-\cite{yosinski2014transferable}).
+\cite{yosinski_how_2014}).
 These techniques observe that features learned in the lower layers of the network are general and transferable, while the top layers are more task-specific.
@@
 
-\section{Experimental setup}
+\section{Experimental setup} \label{sec:experimental_setup}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % add figure of the dataset distribution in file training_lengths.png
-\begin{figure} \label{fig:task_len}
+\begin{figure}
+
     \centering
@@
     \caption{Number of training examples for each task}
-\label{fig:training_lengths}
+    \label{fig:task_len}
 \end{figure}
@@
 
-\section{Results}
+\section{Results} \label{sec:results}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{lstlisting}[gobble=2]
-    \begin{table}[h] 
+\begin{table}[ht] 
     \centering
@@
     \end{table}
-\end{lstlisting}
 
-\begin{figure}%[ht]
+\begin{figure}[ht]
     %\centering
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
     \centering
@@
 
-\section{Discussion}
+\section{Discussion} \label{sec:discussion}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{table}[ht] \label{tbl:task_perplexities}
+\begin{table}[ht] 
     \centering
     \caption{Test perplexity for each task and precision hyperparameter $\tau$}
+    \label{tbl:task_perplexities}
     \resizebox{\linewidth}{!}{%

### 2024-06-13 paper/main.tex
Updates from Overleaf

@@
 
-\title{BoRA: Hierarchical Bayesian Low-Rank Adaption for Multi-task Large Language Models}
+\title{BoRA: Bayesian Hierarchical  Low-Rank Adaption for Multi-task Large Language Models}
 \author[1]{Simen Eide}
@@
 
-BoRA addresses these trade-offs by leveraging a Bayesian hierarchical model that allows tasks to share information through global hierarchical prior parameters.
+BoRA addresses these trade-offs by leveraging a Bayesian hierarchical model that allows tasks to share information through global hierarchical priors.
 This enables tasks with limited data to benefit from the overall structure derived from related tasks while allowing tasks with more data to specialize.
-We demonstrate the effectiveness of BoRA on the next token generation task using the Talk of Norway Dataset, where each parliament representative's speeches represent distinct tasks.
-Our experimental results show that BoRA outperforms both traditional individual and unified model approaches, achieving lower perplexity and better generalization across tasks.
-This method provides a scalable and efficient solution for multi-task LLM finetuning, with significant practical implications for diverse applications such as media content generation, chatbot responses, and scriptwriting.
+Our experimental results show that BoRA outperforms both individual and unified model approaches, achieving lower perplexity and better generalization across tasks.
+This method provides a scalable and efficient solution for multi-task LLM finetuning, with significant practical implications for diverse applications.
 }
@@
 LLMs are typically pretrained on a large and diverse dataset of texts, and then fine-tuned on a new data set of texts, dedicated to a specific task.
-The fine-tuning process is often done by maximizing the next token log likelihood of the task-specific dataset, and the resulting model is then used to make predictions and generations on new data.
-A popular fine-tuning method is Low-Rank Adaption (LoRA), which reduces the number of trainable parameters by a factor of up to 10'000 and the GPU memory requirement by a factor of 3, while still outperforming other finetuning techniques (\cite{hu_lora_2022}).
+The fine-tuning process is often done by maximizing the next token log likelihood of the task-specific dataset, and the resulting model is then used to make predictions and generations for new data.
+A popular fine-tuning method is Low-Rank Adaption (LoRA), which reduces the number of trainable parameters by a factor of up to 10'000 and the GPU memory requirement by a factor of 3, while still outperforming other fine-tuning techniques (\cite{hu_lora_2022}).
 
@@
 A media organization might want to use an LLM to generate various textual elements to generate headlines, keywords, and summaries for a given news article. 
-A website might have multiple chatbot tasks answering questions to their users handling inquiries across a range of topics such as product information, order status, and troubleshooting.
+A website might have multiple chatbot tasks answering user inquiries across various topics.
 And finally, a scriptwriting AI might want to generate dialogues for different characters in a movie, each with their unique style and vocabulary.
@@
 In these multi-task cases, the practitioner can either (1) train one model for each task and dataset separately, or (2) train one model for all tasks and condition the specific output wanted (e.g \cite{Raffel2019}).
-There is a trade-off between these two approaches: Train each task separately and the model will be able to specialize on each task, but it may suffer from limited data available for each task.
+There is a trade-off between these two approaches: Train each task separately and the model will be able to specialize on each task, but it may suffer when limited data are available for some task.
 Train one model for all tasks, and the model will be able to share information between tasks, but it may not be able to specialize on each task due to limited model capacity.
 
-In this paper we suggest a way to circumvent these trade-offs and presents a generalization of the two LoRA fine-tuning methods for multi-task problems, which we call Hierarchical Bayesian Low-Rank Adaption method (BoRA).
-The Hierarchical LoRA method allows tasks to share information between each other through sharing a set of global hierarchical prior parameters, in a Bayesian hierarchical setting. 
-This means that tasks with little data can benefit from the global structure of the related tasks, while tasks with more data can specialize on their specific task.
+In this paper we suggest a way to circumvent these trade-offs and presents a generalization of the two LoRA fine-tuning methods for multi-task problems, which we call Bayesian Hierarchical Low-Rank Adaption method (BoRA).
+The BoRA method allows tasks to share information between each other through sharing a set of global hierarchical prior parameters, in a Bayesian hierarchical setting. 
+This means that tasks with little data can benefit from the global structure across related tasks, while tasks with more data can specialize on their specific task.
 We show in Section \ref{sec:BoRA} that BoRA can be seen as a generalization of the two approaches above.
 
-To test our method, we evaluate the Hierarchical LoRA method on the next token generation problem of parliament speeches letting each parliament representative be a task.
-In our setting, BoRA is superior to both conventional approaches above and suggest a way to mitigate the trade-off between the two.
+To test our method, we evaluate the BoRA on the next token generation problem of parliament speeches letting each parliament representative be a task.
+In our setting, BoRA is superior to both conventional approaches described above and suggests a way to mitigate the trade-off between the two.
 % Write something about how easy it is to implement and how it can be used in practice?
 
-The paper is organized as follows: Section \ref{sec:related_works} presents related works, Section \ref{sec:method} presents the Hierarchical LoRA method, Section \ref{sec:experimental_setup} describes the experimental setup, Section \ref{sec:results} presents the results, and Section \ref{sec:discussion} discusses the results and the implications of the Hierarchical LoRA method.
+The paper is organized as follows: Section \ref{sec:related_works} presents related works.
+Section \ref{sec:method} presents BoRA, the Hierarchical LoRA method; Section \ref{sec:experimental_setup} describes the experimental setup, Section \ref{sec:results} presents the results, and Section \ref{sec:discussion} discusses the results and the implications of the new method.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+We group related work into two areas: multi-task learning and fine-tuning methods of large language models. Autoregressive LLMs and Bayesian modeling are described in Section \ref{sec:method}.
 
@@
 \cite{yosinski_how_2014}).
-These techniques observe that features learned in the lower layers of the network are general and transferable, while the top layers are more task-specific.
+These techniques exploit that features learned in the lower layers of the network are general and transferable, while the top layers are more task-specific.
 This approach has been widely adopted in neural networks.
@@
 There is also the Efficient Low Rank Adaption (LoRA+) \cite{hayou_lora_2024}, which improves the existing LoRA optimization algorithm by adjusting the learning rate of the low-rank parameters.
-All these methods are designed to improve the performance of the LoRA algorithm, and can be used in conjunction with the proposed Hierarchical LoRA method.
+Lastly, MoRA \cite{jiang_mora_2024} decompose the low-rank parameters using learnable square matrices with non-parameter based compressions and decompressions.
+All these methods are designed to improve the performance of language models similar to LoRA, and they can be used in conjunction with the proposed Bayesian Hierarchical LoRA method.
 
@@
 This section describes the Hierarchical LoRA method for multi-task problems. 
-We start by defining the task and data structure, and then introduce the pretrained autoregressive language model.
-We then describe the LoRA method, and finally present our Hierarchical LoRA model.
+We begin by defining the task and data structure, followed by an introduction to pretrained autoregressive language models.
+Next, we describe the LoRA method, and finally, we present the proposed Bayesian Hierarchical LoRA model.
 
 \subsection{Task and Data Structure}
-We define our study across $D$ distinct tasks, denoted as $d=1,2,...,D$.
-For each task $d$, we have a dataset $\mathcal{D}_d$ containing $N_d$ documents, with each document consisting of a sequence of $W_n$ tokens, $n=1,..., N_d$.
+We define our study across $D$ distinct tasks, denoted by $d=1,2,...,D$.
+For each task $d$, we have a dataset $\mathcal{D}_d$ containing $N_d$ documents, with each document $n=1,..., N_d$ consisting of a sequence of $W_n$ tokens.
 This dataset structure can be formally represented as:
@@
 In this notation, $w_{d,n,1:W_n}$ represents the token sequence in the $n^{th}$ document of the $d^{th}$ task.
-For simplification in subsequent discussions, we will omit the indices $d$ and $n$ when their reference is evident from the context.
+For simplification in subsequent discussions, we will omit the indices $d$ and $n$ (and write $w_i$) when their reference is evident from the context.
 
@@
 
-\subsection{The Hierarchical Bayesian LoRA model} \label{sec:BoRA}
+\subsection{The Bayesian Hierarchical LoRA model} \label{sec:BoRA}
 In the hierarchical LoRA model, we introduce a LoRA parameter set for each task $d$, denoted as $\theta_d$, containing all the low-rank decomposition matricies $A_d$ and $B_d$ from each decomposed layer in the Large Language Model.
@@
 $P(\Theta) = 1$).
-The prior assumes that the task parameters share the common origin from the shared Gaussian prior, because the tasks are related: input data come from similar distributions and the output tasks has similar aspects.
+The prior assumes that the task parameters share a common origin from the common Gaussian prior, because the tasks are related: input data come from similar distributions and the output tasks have similar aspects.
 The precision hyperparameter $\tau$ controls how much the task parameters $\theta_d$ are a priori expected to be allowed to deviate from the hierarchical mean parameters $\Theta$. 
-This is essentially a measure for how much structure and information the tasks share with each other, and should be set by the practitioner based on her's prior knowledge.
+This is a measure for how much structure and information the tasks share with each other, and should be set by the practitioner based on her's prior knowledge.
 
@@
 
-The resulting BoRA model balances each task based on its data size and similarity to the other tasks: If a task has few data points, the posterior in equation \ref{eq:posterior} will be dominated by the prior term and the task parameters will be close to the global hierarchical parameters and borrow information and structure from these.
+The resulting BoRA model solves each task based on its data size and similarity to the other tasks: If a task has few data points, the posterior in equation \ref{eq:posterior} will be dominated by the prior term and the task parameters will be close to the global hierarchical parameters and borrow information and structure from these.
 On the other hand, if a task has a lot of data points, the likelihood will dominate the posterior and the task parameters will be allowed to learn more about the specific task and rely less on the hierarchical parameters.
@@
 \subsection{Optimization}
-Given a precision hyperparameter $\tau$, we want to find the maximum a posteriori (MAP) estimate of the hierarchical model. 
+Given the precision hyperparameter $\tau$, we want to find the maximum a posteriori (MAP) estimate of the hierarchical model. 
 We will do this by using AdamW (\cite{adamW}), a gradient-based optimization algorithm, optimizing over the task parameters $\theta_{1:D}$ and the hierarchical mean parameters $\Theta$ using the posterior in equation \ref{eq:posterior}.
@@
     }
-    \caption{Number of training examples for each task}
+    \caption{Number of training examples for each task.}
     \label{fig:task_len}
@@
 % As base model, use opt-350m as it is predominantly trained on English, and only has seen a small amount of Norwegian data from the commoncrawls dataset.
-To test our method, we optimize the model with a range of different precision hyperparameters $\tau$, and evaluate the models' Perplexity on the test set for each task.
+To test our method, we optimize the model with a range of different precision hyperparameters $\tau$, and evaluate the perplexity on the test set for each task.
 % DEF PERPLEXITY
-Perplexity is defined as the exponentiated average negative log-likelihood per token, and is a common metric for evaluating language models. A lower perplexity indicates a better model.
+Perplexity is defined as the exponentiated average negative log-likelihood per token, and is a common metric for evaluating language models.
+A lower perplexity indicates a higher likelihood and a better model.
 We use the Talk of Norway Dataset~\cite{lapponi_talk_2018}, which consists of speeches of Norwegian parliament politicians.
-Each speaker will come from different political parties and geographical areas, but the tasks will still share a common domain. 
+Speakers come from different political parties and geographical areas, but the tasks will still share a common domain, represented by parliamentary speeches in Norway over a short period of time. 
 Therefore, we can consider different speakers of this dataset as different tasks. Specifically, we select chronologically the 25 first speakers that have above 100 speeches in the dataset.
-This gives us a total of 25 tasks with samples ranging from 110 to 477 documents, and an average of 202 documents per task. For some speakers we have significantly less data as for others. 
-The distibution of samples per task can be seen in Figure~\ref{fig:task_len}.
+This gives us a total of 25 tasks with samples ranging from 110 to 477 documents, and an average of 202 documents per task. For some speakers we have significantly less data compared to others. 
+The distribution of samples per task can be seen in Figure~\ref{fig:task_len}.
 We leave 33\% of the data for the test set.
@@
 There are many other task specific factors such as under-represented dialects and political views that can influence the final perplexity.
-Controlling for these factors is difficult and out of scope for this paper, but we believe that these factors may be the reason for this negative result.
+Controlling for these factors is difficult and out of scope for this paper, but we believe that these factors may be the reason for this result.
 
@@
 %% MCMC vs MAP
-In principle, one could bypass the Maximum a Posteriori maximization and adopt a fully Bayesian approach which would tell us how much we can trust the hierarchical model.
-Such an approach  would require the use of Markov Chain Monte Carlo, Variational Inference or other methods.
-This would allow us to sample from the posterior distribution of the hierarchical model, and obtain uncertainty estimates for the task parameters.
+In principle, one could adopt a fully Bayesian approach, instead than the Maximum a Posteriori maximization, which would tell us how much we can trust the hierarchical model.
+Such an approach  would require the use of Markov Chain Monte Carlo (MCMC), Variational Inference or other methods.
+This would require to sample from the posterior distribution of the hierarchical model, and would allow to obtain uncertainty estimates for the task parameters.
 However, MCMC methods are computationally heavy, and it is not clear to us how to best sample from the posterior distribution of the language model in an efficient and scalable way.
-A middle ground between a full bayesian approach and the MAP estimate is to be bayesian in a few parameters.
-For instance, instead of our approach where we maximize the likelihood of the validation set with respect to the precision parameter $\tau$, one could be bayesian and integrate over its empirical posterior distribution obtained within the MCMC.
+A middle ground between a full Bayesian approach and the MAP estimate is to be Bayesian for a few parameters.
+For instance, instead of our approach where we maximize the likelihood of the validation set with respect to the precision parameter $\tau$, one could be Bayesian and integrate over its empirical posterior distribution obtained from the MCMC.
 This would give us an estimate of how well the tasks share information with each other in the specific problem.
 We leave these questions for future work.
+
+\begin{figure}[ht]
+    \centering
+    \resizebox{\linewidth}{!}{%
+    \includegraphics[width=\textwidth]{figures/weights_vs_data_lengths.png}
+    }
+    \caption{The figure shows the L2-distance between each task's adapter weight and the global prior on the y-axis, and the number of training examples on the x-axis.
+We see that, as expected, tasks with more training data  have a larger distance to the global prior.}
+    \label{fig:weights_vs_datalen} 
+\end{figure}
 
@@
 
-% Discuss overall result
-% Fetch perplexity for each task and discuss how it improves results significantly for some tasks (less data?), and less larger data tasks.
-% Give some advice on how to use this method in practice.
-
-\begin{figure}[ht]
-    \centering
-    \resizebox{\linewidth}{!}{%
-    \includegraphics[width=\textwidth]{figures/weights_vs_data_lengths.png}
-    }
-    \caption{The figure shows the L2-distance between each task's adapter weight and the global prior on the y-axis, and the number of training examples on the x-axis.
-We see that, as expected, tasks with more training data will have a larger distance to the global prior.}
-    \label{fig:weights_vs_datalen} 
-\end{figure}
-
 \section{Conclusion}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-This paper presented a generalization of the LoRA fine-tuning method for multi-task problems, which we call the Hierarchical LoRA method.
+This paper presented a new generalization of the LoRA fine-tuning method for multi-task problems, which we call BoRA,  Bayesian Hierarchical Low-Rank Adaption for Multi-task
+Large Language Models.
 The hierarchical model allows tasks to share information between each other through a set of global hierarchical prior parameters, and can be seen as a generalization of the two conventional approaches for multi-task problems.
-We show that the Hierarchical LoRA model can improve the test perplexity for all tasks compared to both training the models independently and training one model for all tasks, and that it can mitigate the trade-off between the two approaches.
+We show that BoRA can improve the test perplexity for all tasks compared to both training the models independently and training one model for all tasks, and that it can mitigate the trade-off between the two approaches.
+We propose to adapt the learning rate to the design parameter that controls the trade off between likelihood and prior.
 Future work includes investigating the effect that BoRA has on more tasks and datasets, investigating how the relative capacity of the global model affects the performance, and investigating the full posterior distribution of the model.
@@
 \section{Task specific test perplexities}
-Table \ref{tbl:task_perplexities} shows the test perplexities for each task when using the hierarchical LoRA model with different precision hyperparameters $\tau$.
+Table \ref{tbl:task_perplexities} shows the test perplexities for each task when using BoRA with different precision hyperparameters $\tau$.
 

### 2024-10-19 paper/main.tex
submission state (pre review changes)

@@
 \documentclass[fullpaper]{nldl}
-\paperID{42}
+\paperID{1}
 \vol{V}
@@
 \affil[1]{University of Oslo}
-\date{March 2024}
+\date{July 2024}
 
-%\usepackage{biblatex} % [backend=biber,style=alphabetic,sorting=ynt] %not necessary when using template
+\usepackage{biblatex}[backend=biber,style=alphabetic,sorting=ynt] %not necessary when using template
 \addbibresource{references.bib}
@@
 
-In these multi-task cases, the practitioner can either (1) train one model for each task and dataset separately, or (2) train one model for all tasks and condition the specific output wanted (e.g \cite{Raffel2019}).
+In these multi-task cases, the practitioner can either (1) train one model for each task and dataset separately, or (2) train one model for all tasks and condition the specific output wanted (e.g~\cite{Raffel2019}).
 There is a trade-off between these two approaches: Train each task separately and the model will be able to specialize on each task, but it may suffer when limited data are available for some task.
@@
 This means that tasks with little data can benefit from the global structure across related tasks, while tasks with more data can specialize on their specific task.
-We show in Section \ref{sec:BoRA} that BoRA can be seen as a generalization of the two approaches above.
+We show in Section~\ref{sec:BoRA} that BoRA can be seen as a generalization of the two approaches above.
 

### 2024-10-19 paper/main.tex
english cleanup

@@
 
-\title{BoRA: Bayesian Hierarchical  Low-Rank Adaption for Multi-task Large Language Models}
+\title{BoRA: Bayesian Hierarchical Low-Rank Adaptation for Multi-task Large Language Models}
 \author[1]{Simen Eide}
@@
 \affil[1]{University of Oslo, Schibsted}
-\affil[1]{University of Oslo}
+\affil[2]{University of Oslo}
 \date{July 2024}
 
-\usepackage{biblatex}[backend=biber,style=alphabetic,sorting=ynt] %not necessary when using template
+\usepackage{biblatex}[backend=biber,style=alphabetic,sorting=ynt] % Not necessary when using template
 \addbibresource{references.bib}
@@
 \maketitle
-\abstract{This paper introduces Bayesian Hierarchical Low-Rank Adaption (BoRA), a novel method for finetuning multi-task Large Language Models (LLMs).
-Current finetuning approaches, such as Low-Rank Adaption (LoRA), perform exeptionally well in reducing training parameters and memory usage but face limitations when applied to multiple similar tasks.
+\abstract{This paper introduces Bayesian Hierarchical Low-Rank Adaptation (BoRA), a novel method for fine-tuning multi-task Large Language Models (LLMs).
+Current fine-tuning approaches, such as Low-Rank Adaptation (LoRA), perform exceptionally well in reducing training parameters and memory usage but face limitations when applied to multiple similar tasks.
 Practitioners usually have to choose between training separate models for each task or a single model for all tasks, both of which come with trade-offs in specialization and data utilization.
@@
 Our experimental results show that BoRA outperforms both individual and unified model approaches, achieving lower perplexity and better generalization across tasks.
-This method provides a scalable and efficient solution for multi-task LLM finetuning, with significant practical implications for diverse applications.
+This method provides a scalable and efficient solution for multi-task LLM fine-tuning, with significant practical implications for diverse applications.
 }
@@
 \section{Introduction}
-Large Language Models (LLM) have become a popular model for solving a wide range of generative text problems.
-LLMs are typically pretrained on a large and diverse dataset of texts, and then fine-tuned on a new data set of texts, dedicated to a specific task.
+Large Language Models (LLMs) have become popular models for solving a wide range of generative text problems.
+LLMs are typically pretrained on large and diverse datasets of texts, and then fine-tuned on new datasets dedicated to specific tasks.
 The fine-tuning process is often done by maximizing the next token log likelihood of the task-specific dataset, and the resulting model is then used to make predictions and generations for new data.
-A popular fine-tuning method is Low-Rank Adaption (LoRA), which reduces the number of trainable parameters by a factor of up to 10'000 and the GPU memory requirement by a factor of 3, while still outperforming other fine-tuning techniques (\cite{hu_lora_2022}).
+A popular fine-tuning method is Low-Rank Adaptation (LoRA), which reduces the number of trainable parameters by a factor of up to 10,000 and the GPU memory requirement by a factor of 3, while still outperforming other fine-tuning techniques (\cite{hu_lora_2022}).
 
-When performing fine-tuning of LLMs, the practitioner often wants to solve multiple similar tasks at the same time: 
-A media organization might want to use an LLM to generate various textual elements to generate headlines, keywords, and summaries for a given news article. 
+When performing fine-tuning of LLMs, the practitioner often wants to solve multiple similar tasks at the same time. For example:
+A media organization might want to use an LLM to generate various textual elements like headlines, keywords, and summaries for a given news article. 
 A website might have multiple chatbot tasks answering user inquiries across various topics.
-And finally, a scriptwriting AI might want to generate dialogues for different characters in a movie, each with their unique style and vocabulary.
+And finally, a scriptwriting AI might want to generate dialogues for different characters in a movie, each with its unique style and vocabulary.
 
-In these multi-task cases, the practitioner can either (1) train one model for each task and dataset separately, or (2) train one model for all tasks and condition the specific output wanted (e.g~\cite{Raffel2019}).
-There is a trade-off between these two approaches: Train each task separately and the model will be able to specialize on each task, but it may suffer when limited data are available for some task.
+In these multi-task cases, the practitioner can either (1) train one model for each task and dataset separately, or (2) train one model for all tasks and condition the specific output wanted (e.g.
+~\cite{Raffel2019}).
+There is a trade-off between these two approaches: Train each task separately and the model will be able to specialize on each task, but it may suffer when limited data are available for some tasks.
 Train one model for all tasks, and the model will be able to share information between tasks, but it may not be able to specialize on each task due to limited model capacity.
 
-In this paper we suggest a way to circumvent these trade-offs and presents a generalization of the two LoRA fine-tuning methods for multi-task problems, which we call Bayesian Hierarchical Low-Rank Adaption method (BoRA).
-The BoRA method allows tasks to share information between each other through sharing a set of global hierarchical prior parameters, in a Bayesian hierarchical setting. 
-This means that tasks with little data can benefit from the global structure across related tasks, while tasks with more data can specialize on their specific task.
+In this paper, we suggest a way to circumvent these trade-offs and present a generalization of the two LoRA fine-tuning methods for multi-task problems, which we call the Bayesian Hierarchical Low-Rank Adaptation method (BoRA).
+The BoRA method allows tasks to share information between each other through a set of global hierarchical prior parameters, in a Bayesian hierarchical setting. 
+This means that tasks with little data can benefit from the global structure across related tasks, while tasks with more data can specialize on their specific tasks.
 We show in Section~\ref{sec:BoRA} that BoRA can be seen as a generalization of the two approaches above.
 
-To test our method, we evaluate the BoRA on the next token generation problem of parliament speeches letting each parliament representative be a task.
+To test our method, we evaluate BoRA on the next token generation problem of parliamentary speeches, letting each parliamentary representative be a task.
 In our setting, BoRA is superior to both conventional approaches described above and suggests a way to mitigate the trade-off between the two.
-% Write something about how easy it is to implement and how it can be used in practice?
-
 The paper is organized as follows: Section \ref{sec:related_works} presents related works.
@@
 
-% Should this fact be added somewhere?
-% If the D tasks are more similar to eachother than the pretraining corpus, the optimal parameters $\hat{\theta}_d$ should share some information during training in order to maximize their likelihood.
+% If the D tasks are more similar to each other than the pretraining corpus, the optimal parameters $\hat{\theta}_d$ should share some information during training in order to maximize their likelihood.
 
-\section{Related works} \label{sec:related_works}
+\section{Related Works} \label{sec:related_works}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Multi-task Learning \cite{caruana_multitask_1997} is a popular machine learning approach for training one model on multiple tasks.
-A common approach is to share the lower layers of a neural network between tasks, and have separate parameters for the last (or last few) layers for each task.
+A common approach is to share the lower layers of a neural network between tasks and have separate parameters for the last (or last few) layers for each task.
 Our paper suggests a different approach, where we share information between tasks through a hierarchical prior over the task-specific parameters.
 
-\subsection{Finetuning methods}
-A popular approach to finetuning neural networks involves adjusting only the top layer (or the n top layers) of a pre-trained neural network (e.g.
-\cite{yosinski_how_2014}).
-These techniques exploit that features learned in the lower layers of the network are general and transferable, while the top layers are more task-specific.
+\subsection{Fine-tuning Methods}
+A popular approach to fine-tuning neural networks involves adjusting only the top layer (or the n top layers) of a pre-trained neural network (e.g., \cite{yosinski_how_2014}).
+These techniques exploit the fact that features learned in the lower layers of the network are general and transferable, while the top layers are more task-specific.
 This approach has been widely adopted in neural networks.
-However, in this paper we focus on the LoRA method, which instead allows us to train a low rank decomposition of the entire network.
+However, in this paper, we focus on the LoRA method, which instead allows us to train a low-rank decomposition of the entire network.
 
@@
 The Sparse Mixture of Low Rank Adapters (SIRA) \cite{zhu_sira_2023}, for instance, creates a Mixture of Expert on the low-rank parameters.
-Another variation, the Weight Decomposed Low-Rank Adaption (DoRA) \cite{hayou_lora_2024}, decomposes the low-rank parameters into scale and unit norm direction parameters.
-There is also the Efficient Low Rank Adaption (LoRA+) \cite{hayou_lora_2024}, which improves the existing LoRA optimization algorithm by adjusting the learning rate of the low-rank parameters.
-Lastly, MoRA \cite{jiang_mora_2024} decompose the low-rank parameters using learnable square matrices with non-parameter based compressions and decompressions.
+Another variation, the Weight Decomposed Low-Rank Adaptation (DoRA) \cite{hayou_lora_2024}, decomposes the low-rank parameters into scale and unit norm direction parameters.
+There is also the Efficient Low Rank Adaptation (LoRA+) \cite{hayou_lora_2024}, which improves the existing LoRA optimization algorithm by adjusting the learning rate of the low-rank parameters.
+Lastly, MoRA \cite{jiang_mora_2024} decomposes the low-rank parameters using learnable square matrices with non-parameter based compressions and decompressions.
 All these methods are designed to improve the performance of language models similar to LoRA, and they can be used in conjunction with the proposed Bayesian Hierarchical LoRA method.
@@
 \subsection{Pretrained Autoregressive Language Model}
-Our foundational model is a pretrained autoregressive language model, denoted by its parameters $\theta_{full}$ and its architecture, referred to as a Large Language Model ($LLM$).
-The parameters $\theta_{full}$ has been obtained from pretraining on a large and diverse dataset, distinct from our current dataset $\mathcal{D}$.
-The model's probability distribution is given by:
+Our foundational model is a pretrained autoregressive language model, characterized by its parameters $\theta_{full}$ and its architecture, referred to as a Large Language Model (LLM).
+The parameters $\theta_{full}$ were acquired through pretraining on a large and diverse dataset, distinct from our current dataset $\mathcal{D}$.
+The model's probability distribution is expressed as:
 
@@
 \end{equation}
-%
-where $LLM(w_{i+1} | w_{1:i})$ is the probability of the next token $w_{i+1}$ given the previous tokens $w_{1:i}$ in the document.
-We omit to write $\theta_{full}$ explicitly in the LLM, as the autoregressive language model always depend on these parameters.
+where $LLM(w_{i+1} | w_{1:i})$ denotes the probability of the next token $w_{i+1}$ given the preceding tokens $w_{1:i}$ in the document.
+For simplicity, we omit $\theta_{full}$ explicitly in the LLM, since the autoregressive language model inherently depends on these parameters.
 
-\subsection{Low-Rank Adaption (LoRA)}
-A Large Language Model contains many linear layers, and the LoRA method reduces the number of trainable parameters by reparameterizing the weight parameters ($W \in \R^{n_1 \times n_2}$, $W \subset \theta$) of these layers into the original pretrained weights $W_{full}$ and a low-rank decomposition $A$ and $B$ (\cite{hayou_lora_2024}):
+\subsection{Low-Rank Adaptation (LoRA)}
+A Large Language Model consists of numerous linear layers.
+The LoRA method reduces the number of trainable parameters by reparameterizing the weight parameters ($W \in \R^{n_1 \times n_2}$, $W \subset \theta$) of these layers into the original pretrained weights $W_{full}$ and a low-rank decomposition $A$ and $B$ (\cite{hayou_lora_2024}):
+
 \begin{equation} \label{eq:LoRA}
@@
 $B \in \R^{n_1 \times r}$ and $A \in \R^{r \times n_2}$ are the low-rank factors, 
-$n_1$, $n_2$ are the input and output dimension of the original weight matrix in the linear layer,
-$r \ll \text{min}(n_1, n_2)$ is the low rank dimension 
+$n_1$ and $n_2$ are the input and output dimensions of the original weight matrix in the linear layer,
+$r \ll \text{min}(n_1, n_2)$ is the low-rank dimension,
 and $\alpha$ is a scaling hyperparameter.
 
-During LoRA fine-tuning, the low-rank factors $A$ and $B$ are learned (estimated) from data, while the original pretrained weights $W_{full}$ are kept fixed to their pretrained value.
+During LoRA fine-tuning, the low-rank factors $A$ and $B$ are learned (estimated) from the data, while the original pretrained weights $W_{full}$ remain fixed to their pretrained values.
 
-\subsection{The Bayesian Hierarchical LoRA model} \label{sec:BoRA}
-In the hierarchical LoRA model, we introduce a LoRA parameter set for each task $d$, denoted as $\theta_d$, containing all the low-rank decomposition matricies $A_d$ and $B_d$ from each decomposed layer in the Large Language Model.
-The likelihood of the data $\mathcal{D}$ given the task parameters $\theta_{1:D}$ is then given by (combining equation \ref{eq:data} and \ref{eq:LLMprob}):
+\subsection{The Bayesian Hierarchical LoRA Model} \label{sec:BoRA}
+In the hierarchical LoRA model, we introduce a LoRA parameter set for each task $d$, noted as $\theta_d$, which contains all the low-rank decomposition matrices $A_d$ and $B_d$ from each decomposed layer in the Large Language Model.
+The likelihood of the dataset $\mathcal{D}$ given the task parameters $\theta_{1:D}$ is then formulated by combining equations \ref{eq:data} and \ref{eq:LLMprob}:
 
@@
 %
-%
-Tasks will share some knowledge with each other, and we model this by introducing a Gaussian hierarchical prior over the task parameters $\theta_{1:D}$, given by
+Tasks share some common knowledge, modeled by introducing a Gaussian hierarchical prior over the task parameters $\theta_{1:D}$, described by:
+
 \begin{equation} \label{eq:prior}
-    P(\theta_{1:D} | \Theta, \tau ) = \prod_{d=1}^D N(\theta_d ; \Theta, \frac{1}{\tau} I)
+    P(\theta_{1:D} | \Theta, \tau) = \prod_{d=1}^D N(\theta_d ; \Theta, \frac{1}{\tau} I)
 \end{equation}
-%
-where $\Theta$ is a set of hierarchical mean parameters, $I$ is a unit diagonal matrix and $\tau \geq 0$ is a scalar precision hyperparameter controlling how similar the task parameters $\theta_d$ should be to the hierarchical mean parameters $\Theta$.
+where $\Theta$ is a set of hierarchical mean parameters, $I$ is a unit diagonal matrix, and $\tau \geq 0$ is a scalar precision hyperparameter controlling how similar the task parameters $\theta_d$ should be to the hierarchical mean parameters $\Theta$.
 We let $\Theta$ have a uniform hyperprior (i.e.
 $P(\Theta) = 1$).
-The prior assumes that the task parameters share a common origin from the common Gaussian prior, because the tasks are related: input data come from similar distributions and the output tasks have similar aspects.
-The precision hyperparameter $\tau$ controls how much the task parameters $\theta_d$ are a priori expected to be allowed to deviate from the hierarchical mean parameters $\Theta$. 
-This is a measure for how much structure and information the tasks share with each other, and should be set by the practitioner based on her's prior knowledge.
+The prior assumes that the task parameters share a common origin because the tasks are related: input data come from similar distributions and the output tasks share similarities.
+The precision hyperparameter $\tau$ dictates how much the task parameters $\theta_d$ are a priori expected to deviate from the hierarchical mean parameters $\Theta$.
+It indicates how much structure and information is shared among tasks and should be set by the practitioner based on prior knowledge.
+The posterior distribution of the hierarchical model, obtained by combining equations \ref{eq:lik} and \ref{eq:prior}, is:
 
-The posterior distribution of the hierarchical model is then given by combining equation \ref{eq:lik} and \ref{eq:prior}:
 \begin{equation} \label{eq:posterior}
@@
 %
-%
+The resulting BoRA model utilizes each task based on its dataset size and similarity to other tasks: If a task has few data points, the posterior in equation \ref{eq:posterior} will be dominated by the prior term, and the task parameters will remain near the global hierarchical parameters, borrowing information from these.
+Conversely, if a task has many data points, the likelihood will dominate the posterior, allowing the task parameters to focus more on the individual task.
 
-The resulting BoRA model solves each task based on its data size and similarity to the other tasks: If a task has few data points, the posterior in equation \ref{eq:posterior} will be dominated by the prior term and the task parameters will be close to the global hierarchical parameters and borrow information and structure from these.
-On the other hand, if a task has a lot of data points, the likelihood will dominate the posterior and the task parameters will be allowed to learn more about the specific task and rely less on the hierarchical parameters.
-
-The hierarchical LoRA model can be seen as a generalization of the two conventional approaches for multi-task problems:
-If we let $\tau$ go to zero, the probability distribution of the prior (eq.
-\ref{eq:prior}) will become a flat uniform prior and the problem will be equivalent to training each task independently.
-On the other hand, if we let $\tau$ go to infinity, the prior will become a strong constraint on the task parameters not allowing them to deviate from the global parameters, and the problem will be equivalent to training one model for all tasks.
+The hierarchical LoRA model can be seen as a generalization of the two conventional approaches for multi-task problems: If we let $\tau$ approach zero, the probability distribution of the prior (eq.
+\ref{eq:prior}) will become a flat uniform prior and the situation will be equivalent to training each task independently.
+On the other hand, if we let $\tau$ approach infinity, the prior will become a strong constraint on the task parameters, preventing them from deviating from the global parameters, and the situation will be equivalent to training one model for all tasks.
 
 \subsection{Optimization}
-Given the precision hyperparameter $\tau$, we want to find the maximum a posteriori (MAP) estimate of the hierarchical model. 
-We will do this by using AdamW (\cite{adamW}), a gradient-based optimization algorithm, optimizing over the task parameters $\theta_{1:D}$ and the hierarchical mean parameters $\Theta$ using the posterior in equation \ref{eq:posterior}.
-This implies that the task parameters with have gradients both towards minimizing the convential LLM loss (represented by the likelihood term) and the hierarchical mean parameters (originating from the prior).
-The hierarchical mean parameters $\Theta$ will have gradients towards the average of the task parameters, and learn to represent the global structure of the tasks.
+Given the precision hyperparameter $\tau$, we seek the maximum a posteriori (MAP) estimate of the hierarchical model.
+We achieve this using AdamW (\cite{adamW}), a gradient-based optimization algorithm, by optimizing over the task parameters $\theta_{1:D}$ and the hierarchical mean parameters $\Theta$ using the posterior in equation \ref{eq:posterior}.
+This implies that the task parameters will have gradients both towards minimizing the conventional LLM loss (represented by the likelihood term) and the hierarchical mean parameters (originating from the prior).
+The hierarchical mean parameters $\Theta$ will have gradients towards the average of the task parameters, learning to represent the global structure of the tasks.
 
 An important observation is that the gradient of the prior term in equation \ref{eq:posterior} is proportional to the precision hyperparameter $\tau$.
-This means that the optimization will need more steps to converge for larger values of $\tau$, as the task parameters will be more constrained towards the hierarchical mean parameters, and a posteriori dependent on each other.
+This means that the optimization will require more steps to converge for larger values of $\tau$, as the task parameters will be more constrained towards the hierarchical mean parameters and a posteriori dependent on each other.
 Therefore, we adjust the learning rate of the AdamW optimizer to be proportional to the precision hyperparameter $\tau$, and we find in our experiments that this causes the optimizations to converge at a similar rate for different values of $\tau$.
@@
 A lower perplexity indicates a higher likelihood and a better model.
-We use the Talk of Norway Dataset~\cite{lapponi_talk_2018}, which consists of speeches of Norwegian parliament politicians.
-Speakers come from different political parties and geographical areas, but the tasks will still share a common domain, represented by parliamentary speeches in Norway over a short period of time. 
-Therefore, we can consider different speakers of this dataset as different tasks. Specifically, we select chronologically the 25 first speakers that have above 100 speeches in the dataset.
-This gives us a total of 25 tasks with samples ranging from 110 to 477 documents, and an average of 202 documents per task. For some speakers we have significantly less data compared to others. 
+We use the Talk of Norway Dataset~\cite{lapponi_talk_2018}, which consists of speeches by Norwegian parliament politicians.
+Speakers come from different political parties and geographical areas, but the tasks will still share a common domain represented by parliamentary speeches in Norway over a short period of time.
+Therefore, we can consider different speakers of this dataset as different tasks.
+Specifically, we select the first 25 speakers chronologically who have more than 100 speeches in the dataset.
+This results in 25 tasks with samples ranging from 110 to 477 documents, and an average of 202 documents per task.
+For some speakers, we have significantly less data compared to others.
 The distribution of samples per task can be seen in Figure~\ref{fig:task_len}.
-We leave 33\% of the data for the test set.
+We reserve 33\% of the data for the test set.
 
 We select a model that has been pretrained on a large and diverse dataset, but has not seen the specific tasks that we want to evaluate it on.
-We will use the `opt-350m' model~\cite{zhang_opt_2022}, which is predominantly trained on English, and has only seen a small amount of Norwegian data from the CommonCrawls dataset.
-
+We use the `opt-350m' model~\cite{zhang_opt_2022}, which is predominantly trained on English, and has only seen a small amount of Norwegian data from the CommonCrawls dataset.
 \section{Results} \label{sec:results}
@@
 We see that the model achieves the best test perplexity of 12.82 when using a precision parameter of $100$.
-The improvement is largest from training the models independently ($\tau = 0$), and smaller when comparing to the limiting case when all task specific model parameters are constraint to be equal ($\tau=10000$).
+The improvement is largest compared to training the models independently ($\tau = 0$), and smaller when comparing to the limiting case where all task-specific model parameters are constrained to be equal ($\tau=10000$).
 
@@
     }
-        \caption{Empirical results of the hierarchical LORA finetuned on the three tasks. The perplexity is calculated on the test set.}
+        \caption{Empirical results of the hierarchical LoRA fine-tuned on the three tasks. The perplexity is calculated on the test set.}
         \label{tbl:results}
@@
     }
-    \caption{Plot of Precision vs Perplexity. The thick black line is the overall test perplexity over all tasks, and the thinner grey lines are the test perplexity for each individual task.}
+    \caption{Plot of Precision vs Perplexity. The thick black line is the overall test perplexity across all tasks, and the thinner grey lines represent the test perplexity for each individual task.}
     \label{fig:results_plot}
@@
 
-If we consider each task separately (Table \ref{tbl:task_perplexities}), we see that all tasks benefit from a hierarchical model compared to both the case when all models are trained independently ($\tau = 0$) and the limiting case when all task specific model parameters are constraint to be equal ($\tau=10000$).
-Figure \ref{fig:relative_improvement} shows the relative improvement of each task when using the hierarchical model compared to the two cases versus their training data size.
-There is an indication that the improvement from training separately is larger for tasks with less data, and we do not see this for the "one model" case.
+When considering each task separately (Table \ref{tbl:task_perplexities}), we see that all tasks benefit from a hierarchical model compared to both the case when all models are trained independently ($\tau = 0$) and the limiting case where all task-specific model parameters are constrained to be equal ($\tau=10000$).
+Figure \ref{fig:relative_improvement} shows the relative improvement of each task when using the hierarchical model compared to the two cases versus their dataset sizes.
+There is an indication that the improvement from training separately is larger for tasks with less data, while this is not observed for the "one model" case.
 A precision parameter of $100$ gives the best perplexity for all our tasks.
-
 \begin{figure}[ht]
@@
     }
-    \caption{Task dataset size versus the relative improvement in perplexity of the best performing hierarchical model ($\tau = 100$) compared to the case when all models are trained independently ($\tau = 0$) in blue, and compared to the limiting case when all task specific model parameters are constraint to be equal ($\tau=10000$) in red.
-    The figure shows that all tasks benefit from sharing parameters with the global model, and indicates that the tasks with less data benefit more than the tasks with more data.}
+    \caption{Task dataset size versus the relative improvement in perplexity of the best performing hierarchical model ($\tau = 100$) compared to the case when all models are trained independently ($\tau = 0$) in blue, and compared to the limiting case when all task-specific model parameters are constrained to be equal ($\tau=10000$) in red.
+The figure shows that all tasks benefit from sharing parameters with the global model, and indicates that tasks with less data benefit more than those with more data.}
     \label{fig:relative_improvement}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Our results in Figure \ref{fig:results_plot} show that the hierarchical LoRA model can improve the test perplexity for all tasks compared to both training the models independently and training one model for all tasks.
-This confirms our hypothesis that BoRA can be a useful method for multi-task problems, and that it can mitigate the trade-off between training each task separately and training one model for all tasks.
-It is interesting to note the BoRA also helps for the tasks with relatively more data.
+Our results in Figure \ref{fig:results_plot} demonstrate that the hierarchical LoRA model can improve test perplexity for all tasks compared to both training the models independently and training one model for all tasks.
+This confirms our hypothesis that BoRA can be a useful method for multi-task problems, mitigating the trade-off between training each task separately and training one model for all tasks.
+Notably, BoRA also benefits tasks with relatively more data.
+We observe in Figure \ref{fig:weights_vs_datalen} that the distance between the task parameters and the global prior increases with increasing task dataset size.
+This is expected in a hierarchical model, as tasks with more data will have a larger likelihood term in the posterior, allowing them to specialize more on their specific task during training.
 
-We observe in Figure \ref{fig:weights_vs_datalen} that the distance between the task parameters and the global prior increases with increasing task dataset size.
-This is as expected in a hierarchical model, as tasks with more data will have a larger likelihood term in the posterior and will be allowed to specialize more on their specific task during training.
-
-We find no relationship between dataset size and the final perplexity (Figure \ref{fig:absolute_improvement}).
-One could expect that tasks with more data are able to learn more about their specific task and achieve a lower perplexity, but this is not the case in our experiments.
-There are many other task specific factors such as under-represented dialects and political views that can influence the final perplexity.
-Controlling for these factors is difficult and out of scope for this paper, but we believe that these factors may be the reason for this result.
-
-However, when considering the relative improvement of the best performing model, the results in Figure \ref{fig:relative_improvement} show tasks with less data may benefit more than the tasks with more data.
-Although a weak signal, this is an expected result, as the hierarchical model allows tasks with little data to borrow information and structure from the global model, while tasks with more data can specialize on their specific task.
+We observe no consistent relationship between dataset size and the final perplexity (Figure \ref{fig:absolute_improvement}).
+Although one might expect that tasks with more data would learn more about their specific task and achieve lower perplexity, this pattern is not evident in our experiments.
+Other task-specific factors, such as under-represented dialects and political views, might influence the final perplexity.
+Controlling for these factors is complex and beyond the scope of this paper, but they are likely contributors to this outcome.
+However, when considering the relative improvement of our best-performing model, the results in Figure \ref{fig:relative_improvement} indicate that tasks with less data may benefit more than those with more data.
+Although the signal is subtle, this is an expected result, as the hierarchical model allows tasks with limited data to leverage information and structure from the global model, while tasks with more data can specialize in their specific task.
 
 %% MCMC vs MAP
-In principle, one could adopt a fully Bayesian approach, instead than the Maximum a Posteriori maximization, which would tell us how much we can trust the hierarchical model.
-Such an approach  would require the use of Markov Chain Monte Carlo (MCMC), Variational Inference or other methods.
-This would require to sample from the posterior distribution of the hierarchical model, and would allow to obtain uncertainty estimates for the task parameters.
-However, MCMC methods are computationally heavy, and it is not clear to us how to best sample from the posterior distribution of the language model in an efficient and scalable way.
-A middle ground between a full Bayesian approach and the MAP estimate is to be Bayesian for a few parameters.
-For instance, instead of our approach where we maximize the likelihood of the validation set with respect to the precision parameter $\tau$, one could be Bayesian and integrate over its empirical posterior distribution obtained from the MCMC.
-This would give us an estimate of how well the tasks share information with each other in the specific problem.
-We leave these questions for future work.
+In principle, one could adopt a fully Bayesian approach, instead of Maximum a Posteriori (MAP) maximization, to assess the reliability of the hierarchical model.
+Such an approach would necessitate the use of Markov Chain Monte Carlo (MCMC), Variational Inference, or other methods.
+
+This approach would involve sampling from the posterior distribution of the hierarchical model, allowing for the estimation of uncertainty in the task parameters.
+However, MCMC methods are computationally intensive, and it remains unclear to us how to efficiently and scalably sample from the posterior distribution of the language model.
+A middle ground between a full Bayesian approach and the MAP estimate could involve being Bayesian for select parameters.
+For example, instead of our current approach where we maximize the likelihood of the validation set concerning the precision parameter $\tau$, one could be Bayesian and integrate over its empirical posterior distribution obtained from MCMC.
+This would provide insights into how well tasks share information in a specific problem context.
+We leave these explorations for future research.
 
@@
     }
-    \caption{The figure shows the L2-distance between each task's adapter weight and the global prior on the y-axis, and the number of training examples on the x-axis.
-We see that, as expected, tasks with more training data  have a larger distance to the global prior.}
+    \caption{The figure shows the L2-distance between each task's adapter weights and the global prior on the y-axis, and the number of training examples on the x-axis.
+As expected, tasks with more training data have a larger distance to the global prior.}
     \label{fig:weights_vs_datalen} 
@@
     }
-    \caption{Training size vs perplexity of the best performing hierarchical model ($\tau = 100$). The figure shows that there is no strong relationship between dataset size and final perplexity.}
+    \caption{Training size vs. perplexity of the best-performing hierarchical model ($\tau = 100$). The figure shows that there is no strong relationship between dataset size and final perplexity.}
     \label{fig:absolute_improvement}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-This paper presented a new generalization of the LoRA fine-tuning method for multi-task problems, which we call BoRA,  Bayesian Hierarchical Low-Rank Adaption for Multi-task
-Large Language Models.
-The hierarchical model allows tasks to share information between each other through a set of global hierarchical prior parameters, and can be seen as a generalization of the two conventional approaches for multi-task problems.
-We show that BoRA can improve the test perplexity for all tasks compared to both training the models independently and training one model for all tasks, and that it can mitigate the trade-off between the two approaches.
-We propose to adapt the learning rate to the design parameter that controls the trade off between likelihood and prior.
-Future work includes investigating the effect that BoRA has on more tasks and datasets, investigating how the relative capacity of the global model affects the performance, and investigating the full posterior distribution of the model.
+This paper presents a novel generalization of the LoRA fine-tuning method for multi-task problems, which we call BoRA, Bayesian Hierarchical Low-Rank Adaptation for Multi-task Large Language Models.
+The hierarchical model allows tasks to share information between each other through a set of global hierarchical prior parameters and can be seen as a generalization of the two conventional approaches for multi-task problems.
+We demonstrate that BoRA can improve test perplexity for all tasks compared to both training the models independently and training a single model for all tasks, effectively mitigating the trade-off between these two approaches.
+We propose adjusting the learning rate according to the design parameter that controls the trade-off between likelihood and prior.
+Future work includes investigating the effect of BoRA on more tasks and datasets, examining how the relative capacity of the global model affects performance, and exploring the full posterior distribution of the model.
 
@@
 \appendix
-\section{Task specific test perplexities}
+\section{Task-Specific Test Perplexities}
 Table \ref{tbl:task_perplexities} shows the test perplexities for each task when using BoRA with different precision hyperparameters $\tau$.

### 2024-10-19 paper/main.tex
create some newlines to avoid latex long line error
updates on reference++

@@
 The fine-tuning process is often done by maximizing the next token log likelihood of the task-specific dataset, and the resulting model is then used to make predictions and generations for new data. 
-A popular fine-tuning method is Low-Rank Adaptation (LoRA), which reduces the number of trainable parameters by a factor of up to 10,000 and the GPU memory requirement by a factor of 3, while still outperforming other fine-tuning techniques (\cite{hu_lora_2022}).
-
-When performing fine-tuning of LLMs, the practitioner often wants to solve multiple similar tasks at the same time. For example:
+A popular fine-tuning method is Low-Rank Adaptation (LoRA), which reduces the number of trainable parameters by a factor of up to 10,000 and the GPU memory requirement by a factor of 3, while still outperforming other fine-tuning techniques (\cite{huLoRALowRankAdaptation2022}).
+When performing fine-tuning of LLMs, the practitioner often wants to solve multiple similar tasks at the same time. 
+For example:
 A media organization might want to use an LLM to generate various textual elements like headlines, keywords, and summaries for a given news article. 
@@
 
-In these multi-task cases, the practitioner can either (1) train one model for each task and dataset separately, or (2) train one model for all tasks and condition the specific output wanted (e.g.
-~\cite{Raffel2019}).
+In these multi-task cases, the practitioner can either (1) train one model for each task and dataset separately, or (2) train one model for all tasks and condition the specific output wanted (e.g.~\cite{Raffel2019}).
 There is a trade-off between these two approaches: Train each task separately and the model will be able to specialize on each task, but it may suffer when limited data are available for some tasks. 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-We group related work into two areas: multi-task learning and fine-tuning methods of large language models. Autoregressive LLMs and Bayesian modeling are described in Section \ref{sec:method}.
+We group related work into two areas: multi-task learning and fine-tuning methods of large language models. 
+Autoregressive LLMs and Bayesian modeling are described in Section \ref{sec:method}.
 
 \subsection{Multi-task Learning}
-Multi-task Learning \cite{caruana_multitask_1997} is a popular machine learning approach for training one model on multiple tasks.
+Multi-task Learning \cite{caruanaMultitaskLearning1997} is a popular machine learning approach for training one model on multiple tasks. 
 A common approach is to share the lower layers of a neural network between tasks and have separate parameters for the last (or last few) layers for each task. 
 Our paper suggests a different approach, where we share information between tasks through a hierarchical prior over the task-specific parameters.
-
 \subsection{Fine-tuning Methods}
-A popular approach to fine-tuning neural networks involves adjusting only the top layer (or the n top layers) of a pre-trained neural network (e.g., \cite{yosinski_how_2014}).
+A popular approach to fine-tuning neural networks involves adjusting only the top layer (or the n top layers) of a pre-trained neural network (e.g., \cite{yosinskiHowTransferableAre2014}). 
 These techniques exploit the fact that features learned in the lower layers of the network are general and transferable, while the top layers are more task-specific. 
@@
 There are several variations of the LoRA algorithm. 
-The Sparse Mixture of Low Rank Adapters (SIRA) \cite{zhu_sira_2023}, for instance, creates a Mixture of Expert on the low-rank parameters.
-Another variation, the Weight Decomposed Low-Rank Adaptation (DoRA) \cite{hayou_lora_2024}, decomposes the low-rank parameters into scale and unit norm direction parameters.
-There is also the Efficient Low Rank Adaptation (LoRA+) \cite{hayou_lora_2024}, which improves the existing LoRA optimization algorithm by adjusting the learning rate of the low-rank parameters.
-Lastly, MoRA \cite{jiang_mora_2024} decomposes the low-rank parameters using learnable square matrices with non-parameter based compressions and decompressions.
+The Sparse Mixture of Low Rank Adapters (SIRA) \cite{zhuSiRASparseMixture2023}, for instance, creates a Mixture of Expert on the low-rank parameters. 
+Another variation, the Weight Decomposed Low-Rank Adaptation (DoRA) \cite{hayouLoRAEfficientLow2024}, decomposes the low-rank parameters into scale and unit norm direction parameters. 
+There is also the Efficient Low Rank Adaptation (LoRA+) \cite{hayouLoRAEfficientLow2024}, which improves the existing LoRA optimization algorithm by adjusting the learning rate of the low-rank parameters.
+Lastly, MoRA \cite{jiangMoRAHighRankUpdating2024} decomposes the low-rank parameters using learnable square matrices with non-parameter based compressions and decompressions.
 All these methods are designed to improve the performance of language models similar to LoRA, and they can be used in conjunction with the proposed Bayesian Hierarchical LoRA method.
@@
 A Large Language Model consists of numerous linear layers.
-The LoRA method reduces the number of trainable parameters by reparameterizing the weight parameters ($W \in \R^{n_1 \times n_2}$, $W \subset \theta$) of these layers into the original pretrained weights $W_{full}$ and a low-rank decomposition $A$ and $B$ (\cite{hayou_lora_2024}):
+The LoRA method reduces the number of trainable parameters by reparameterizing the weight parameters ($W \in \R^{n_1 \times n_2}$, $W \subset \theta$) of these layers into the original pretrained weights $W_{full}$ and a low-rank decomposition $A$ and $B$ (\cite{hayouLoRAEfficientLow2024}):
 
@@
 Given the precision hyperparameter $\tau$, we seek the maximum a posteriori (MAP) estimate of the hierarchical model.
-We achieve this using AdamW (\cite{adamW}), a gradient-based optimization algorithm, by optimizing over the task parameters $\theta_{1:D}$ and the hierarchical mean parameters $\Theta$ using the posterior in equation \ref{eq:posterior}.
+We achieve this using AdamW (\cite{loshchilovDecoupledWeightDecay2018}), a gradient-based optimization algorithm, by optimizing over the task parameters $\theta_{1:D}$ and the hierarchical mean parameters $\Theta$ using the posterior in equation \ref{eq:posterior}.
 This implies that the task parameters will have gradients both towards minimizing the conventional LLM loss (represented by the likelihood term) and the hierarchical mean parameters (originating from the prior).
@@
 A lower perplexity indicates a higher likelihood and a better model. 
-We use the Talk of Norway Dataset~\cite{lapponi_talk_2018}, which consists of speeches by Norwegian parliament politicians.
+We use the Talk of Norway Dataset~\cite{lapponiTalkNorwayRichly2018}, which consists of speeches by Norwegian parliament politicians. 
 Speakers come from different political parties and geographical areas, but the tasks will still share a common domain represented by parliamentary speeches in Norway over a short period of time. 
@@
 Specifically, we select the first 25 speakers chronologically who have more than 100 speeches in the dataset. 
-This results in 25 tasks with samples ranging from 110 to 477 documents, and an average of 202 documents per task.
-For some speakers, we have significantly less data compared to others.
+This results in 25 tasks with samples ranging from 110 to 477 documents, and an average of 202 documents per task. For some speakers, we have significantly less data compared to others. 
 The distribution of samples per task can be seen in Figure~\ref{fig:task_len}. 
@@
 We select a model that has been pretrained on a large and diverse dataset, but has not seen the specific tasks that we want to evaluate it on.
-We use the `opt-350m' model~\cite{zhang_opt_2022}, which is predominantly trained on English, and has only seen a small amount of Norwegian data from the CommonCrawls dataset.
+We use the `opt-350m' model~\cite{zhangOPTOpenPretrained2022}, which is predominantly trained on English, and has only seen a small amount of Norwegian data from the CommonCrawls dataset.
 \section{Results} \label{sec:results}
@@
     }
-    \caption{Plot of Precision vs Perplexity. The thick black line is the overall test perplexity across all tasks, and the thinner grey lines represent the test perplexity for each individual task.}
+    \caption{Plot of Precision vs Perplexity.
+The thick black line is the overall test perplexity across all tasks, and the thinner grey lines represent the test perplexity for each individual task.
+The leftmost point corresponds to training each task independently ($\tau = 0$), and the rightmost point corresponds to the limiting case when all task-specific model parameters are constrained to be equal ($\tau \rightarrow \infty$).}
     \label{fig:results_plot}

### 2024-12-03 paper/main.tex
add some related literature

@@
 %\documentclass{article}
-\documentclass[fullpaper]{nldl}
+\documentclass[fullpaper,final]{nldl}
 \paperID{1}
@@
 \date{July 2024}
+\vol{265}
 
 \usepackage{biblatex}[backend=biber,style=alphabetic,sorting=ynt] % Not necessary when using template
-\addbibresource{references.bib}
+\addbibresource{My Library.bib}
 \usepackage{amsmath}
@@
 Section \ref{sec:method} presents BoRA, the Hierarchical LoRA method; Section \ref{sec:experimental_setup} describes the experimental setup, Section \ref{sec:results} presents the results, and Section \ref{sec:discussion} discusses the results and the implications of the new method.
+Finally, we make the code openly available at \url{https://github.com/simeneide/bora}.
 
@@
 Our paper suggests a different approach, where we share information between tasks through a hierarchical prior over the task-specific parameters.
+
+Modern training of language models can itself be considered a form of multi-task learning where the model is trained on a diverse set of tasks and datasets \cite{raffelExploringLimitsTransfer2023,brownLanguageModelsAre2020}.
+However, with the trend of even bigger models, pretraining of LLMs is only available to a few large organizations due to both computational and data requirements.
+Therefore, there is a need for methods that can leverage the pretraining of these models for smaller tasks and datasets such as the one we present in this paper.
+
+Model merging of LLMs \cite{yadavTIESMergingResolvingInterference2023,yuLanguageModelsAre2024,liDeepModelFusion2023} is another popular approach, often showing promising performance.
+It takes multiple task-specific models and merges them into one model.
+The standard way is by doing a weighted average of the model parameters (\cite{liDeepModelFusion2023}), but more sophisticated methods exist (\cite{yadavTIESMergingResolvingInterference2023, yuLanguageModelsAre2024}).
+Model merging differs from our approach in that it merges the models after training, while the BoRA method shares information between tasks during training through a Bayesian hierarchical model."
+
 \subsection{Fine-tuning Methods}
@@
 where $\Theta$ is a set of hierarchical mean parameters, $I$ is a unit diagonal matrix, and $\tau \geq 0$ is a scalar precision hyperparameter controlling how similar the task parameters $\theta_d$ should be to the hierarchical mean parameters $\Theta$.
-We let $\Theta$ have a uniform hyperprior (i.e.
-$P(\Theta) = 1$).
+We let $\Theta$ have an improper uniform hyperprior (i.e.
+$P(\Theta) = 1$ for all values of $\Theta$).
 The prior assumes that the task parameters share a common origin because the tasks are related: input data come from similar distributions and the output tasks share similarities.
